This document shows the code for the Exploratory Data Anlysis that will be done using the census demographic dataset.  

## Load necessary libraries

```{r}
# Install necessary libraries 
install.packages(c("dplyr", "ggplot2", "sf"))

# Load libraries
library(dplyr)
library(ggplot2)
library(sf)
library(readr)
```

## Load census data

```{r}
# Load the census demographic dataset
census_data <- read_csv("/Users/anamariasteinercorrea/Documents/thesis/census_data.csv")
msoa_geometry <- st_read("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/MSOA_2011_London_gen_MHW.shp")
```

## Adjust and correct data

```{r}
# View column names and types
str(census_data)

# Get a summary of key variables
summary(census_data)

# Check for missing values
sapply(census_data, function(x) sum(is.na(x)))
```

```{r}
# Remove rows with any NA values
census_data_clean <- na.omit(census_data)
```

```{r}
# Check if the dataset is an sf object
class(census_data_clean)

# Join to msoa shapefile
library(dplyr)
msoa_census <- msoa_geometry %>%
    left_join(census_data_clean, by = "MSOA11NM", "MSOA11CD")

colnames(msoa_census)

# Clean final columns
msoa_census_final <- msoa_census[, c("MSOA11CD.x", "MSOA11NM", "LAD11NM", "Year", "Nationality", "Proportion", "Dissimilarity_Index", "lagged_diss", "H", "Isolation_Index", "geometry.x")]
```

```{r}
# Calculate statistics
min_value <- min(msoa_census_final$Diss_Index, na.rm = TRUE)
max_value <- max(msoa_census_final$Diss_Index, na.rm = TRUE)
median_value <- median(msoa_census_final$Diss_Index, na.rm = TRUE)
mean_value <- mean(msoa_census_final$Diss_Index, na.rm = TRUE)

# Print results
cat("Min:", min_value, "\n")
cat("Max:", max_value, "\n")
cat("Median:", median_value, "\n")
cat("Mean:", mean_value, "\n")
```

## EDA

### Visualice histograms and boxplots of distributions
```{r}
# Directory for histogram and boxplot outputs
dir.create("distribution_plots", showWarnings = FALSE)

# Histograms for all indices
for (var in c("Proportion", "Diss_Index", "Isolation", "Theil_H")) {
  hist_plot <- ggplot(msoa_census_final, aes_string(x = var)) +
    geom_histogram(bins = 30, fill = "blue", color = "black", alpha = 0.7) +
    facet_wrap(~ Nationality, scales = "free") +
    labs(title = paste("Histogram of", var, "by Nationality"),
         x = var, y = "Count") +
    theme_minimal()
  
  # Save histogram
  ggsave(filename = paste0("distribution_plots/Histogram_", var, ".png"),
         plot = hist_plot, width = 10, height = 6)
}

# Boxplots for all indices
for (var in c("Proportion", "Diss_Index", "Isolation", "Theil_H")) {
  box_plot <- ggplot(msoa_census_final, aes(x = Nationality, y = !!sym(var))) +
    geom_boxplot(fill = "lightblue", color = "black", alpha = 0.7) +
    labs(title = paste("Boxplot of", var, "by Nationality"),
         x = "Nationality", y = var) +
    theme_minimal()
  
  # Save boxplot
  ggsave(filename = paste0("distribution_plots/Boxplot_", var, ".png"),
         plot = box_plot, width = 10, height = 6)
}
```

### Visualice temporal line graphs of all indices
```{r}
library(ggplot2)
library(dplyr)

# Ensure Year is numeric
msoa_census_final$Year <- as.numeric(msoa_census_final$Year)

# Aggregate data by Year and Nationality
aggregated_data <- msoa_census_final %>%
  group_by(Year, Nationality) %>%
  summarize(
    Proportion = mean(Proportion, na.rm = TRUE),
    Diss_Index = mean(Diss_Index, na.rm = TRUE),
    Lagged_Diss = mean(lagged_diss, na.rm = TRUE),
    Theil_H = mean(Theil_H, na.rm = TRUE),
    Isolation = mean(Isolation, na.rm = TRUE),
    .groups = "drop"
  )

# Create directory for nationality-wise plots
dir.create("line_graphs_nationality", showWarnings = FALSE)

# Loop through each nationality and create line plots
nationalities <- unique(aggregated_data$Nationality)

for (nationality in nationalities) {
  # Filter data for the current nationality
  nat_data <- aggregated_data %>%
    filter(Nationality == nationality)

  # Generate the line plot
  plot <- ggplot(nat_data) +
    geom_line(aes(x = Year, y = Proportion, color = "Proportion"), size = 1.2) +
    geom_line(aes(x = Year, y = Diss_Index, color = "Dissimilarity Index"), size = 1.2) +
    geom_line(aes(x = Year, y = Theil_H, color = "Theil's H"), size = 1.2) +
    geom_line(aes(x = Year, y = Isolation, color = "Isolation Index"), size = 1.2) +
    labs(
      title = paste("Temporal Trends for", nationality),
      x = "Year", y = "Value", color = "Index"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      legend.position = "top"
    )

  # Save the plot
  ggsave(
    filename = paste0("line_graphs_nationality/Trends_", nationality, ".png"),
    plot = plot,
    width = 10,
    height = 6
  )
}
```

### Visualice and compare temporal line graph between all nationalities
```{r}
# Create line graph for Dissimilarity Index comparing all nationalities
dissimilarity_plot <- ggplot(aggregated_data, aes(x = Year, y = Diss_Index, color = Nationality, group = Nationality)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(title = "Dissimilarity Index Over Time (All Nationalities)",
       x = "Year", y = "Dissimilarity Index", color = "Nationality") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

# Save the plot
dir.create("line_graphs_comparison", showWarnings = FALSE)
ggsave(filename = "line_graphs_comparison/Dissimilarity_Index_All_Nationalities.png",
       plot = dissimilarity_plot, width = 10, height = 6)
```

### Calculate spatial autocorrelation of diss index between nationalities

```{r}
# Required Libraries
library(sf)
library(dplyr)
library(ggplot2)
library(Matrix)
library(FNN)

# Create unique geometries with MSOA codes
unique_geometries <- msoa_census_final %>%
  group_by(MSOA11CD.x) %>%
  summarize(geometry = first(geometry)) %>%
  st_as_sf()

# Ensure consistent CRS for geometries
unique_geometries <- st_transform(unique_geometries, crs = 27700)  # Example: British National Grid

# Loop through each nationality
nationalities <- unique(msoa_census_final$Nationality)

for (nationality in nationalities) {
  # Filter data for the current nationality
  nat_data <- msoa_census_final %>%
    filter(Nationality == nationality) %>%
    select(MSOA11CD.x, Diss_Index) %>%
    distinct()

  # Use st_join for spatial join (ensuring proper handling of sf objects)
  nat_data <- st_join(unique_geometries, nat_data, left = TRUE)

  # Ensure nat_data is still an sf object
  nat_data <- st_as_sf(nat_data)

  # Prepare dissimilarity index values and handle missing data
  x <- nat_data$Diss_Index
  x[is.na(x)] <- mean(x, na.rm = TRUE)  # Replace NA with mean value
  x_mean <- mean(x, na.rm = TRUE)

  # Recompute centroids and weights matrix for this subset
  coords <- st_coordinates(st_centroid(st_geometry(nat_data)))
  knn <- get.knn(coords, k = 10)

  # Create a sparse weights matrix for this subset
  weights <- sparseMatrix(
    i = rep(1:nrow(knn$nn.index), each = 10),
    j = as.vector(knn$nn.index),
    x = 1,
    dims = c(nrow(coords), nrow(coords))
  )

  # Row-standardize the weights matrix
  row_sums <- rowSums(weights)
  row_standardized_weights <- weights / row_sums

  # Compute Local Moran's I manually
  local_moran <- numeric(length(x))
  for (i in seq_along(x)) {
    neighbors <- which(row_standardized_weights[i, ] > 0)
    lagged_value <- sum(row_standardized_weights[i, neighbors] * x[neighbors], na.rm = TRUE)
    local_moran[i] <- (x[i] - x_mean) * (lagged_value - x_mean)
  }

  # Normalize Local Moran's I
  local_moran <- local_moran / var(x, na.rm = TRUE)

  # Add Moran's I results to nat_data
  nat_data <- nat_data %>%
    mutate(Local_I = local_moran,
           Quadrant = case_when(
             (Diss_Index > x_mean & Local_I > 0) ~ "High-High",
             (Diss_Index < x_mean & Local_I > 0) ~ "Low-Low",
             (Diss_Index > x_mean & Local_I < 0) ~ "High-Low",
             (Diss_Index < x_mean & Local_I < 0) ~ "Low-High",
             TRUE ~ "Non-significant"
           ))

  # Create LISA Map
  lisa_map <- ggplot(data = nat_data) +
    geom_sf(aes(fill = Quadrant), color = NA) +
    scale_fill_manual(values = c("High-High" = "red", "Low-Low" = "blue", 
                                 "High-Low" = "orange", "Low-High" = "green", 
                                 "Non-significant" = "grey")) +
    labs(title = paste("LISA Map for", nationality),
         fill = "Quadrant") +
    theme_minimal()

  # Save LISA Map
  if (!dir.exists("moran_outputs")) dir.create("moran_outputs")
  ggsave(filename = paste0("moran_outputs/LISA_Map_", nationality, ".png"),
         plot = lisa_map, width = 10, height = 8)

  # Print progress
  print(paste("Completed Moran's I and LISA Map for:", nationality))
}
```

```{r}
# Moran's I	Meaning
> 0	Positive spatial autocorrelation (clustering)
< 0	Negative spatial autocorrelation (dispersion)
≈ 0	No spatial autocorrelation (random)
P-Value	Interpretation
< 0.05	Significant spatial autocorrelation
> 0.05	No significant spatial autocorrelation
```

```{r}
Mixed Red/Orange Areas

Likely represent historically segregated regions.
These regions may be where immigrant communities settled initially, creating dense cultural and ethnic clusters.
These areas may require policies aimed at fostering integration, such as:
Affordable mixed-income housing.
Educational investments to promote mobility.
Mixed Blue/Green Areas

Likely represent integrated suburban or redevelopment zones.
These areas might reflect a balance of diverse groups, driven by:
Suburban housing policies.
Urban regeneration and investment in infrastructure.
These regions might provide insights into successful policies for replication in other areas.
```

### Visualice maps of diss index between nationalities and through time
```{r}
library(ggplot2)
library(sf)

# Ensure the dataset is an sf object
if (!st_is_sf(msoa_census_final)) {
  msoa_census_final <- st_as_sf(msoa_census_final)
}

# Create directory for spatial maps
dir.create("spatial_maps", showWarnings = FALSE)

# Loop through each nationality to create maps
nationalities <- unique(msoa_census_final$Nationality)

for (nationality in nationalities) {
  # Filter data for the current nationality
  nat_data <- msoa_census_final %>%
    filter(Nationality == nationality)
  
  # Plot map for the current nationality
  diss_map <- ggplot(nat_data) +
    geom_sf(aes(fill = Diss_Index, geometry = geometry), color = NA) +
    scale_fill_viridis_c(option = "C", name = "Dissimilarity Index") +
    labs(
      title = paste("Dissimilarity Index Across London -", nationality),
      caption = "Source: Census Data"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      legend.position = "right"
    )
  
  # Save the plot
  ggsave(
    filename = paste0("spatial_maps/Dissimilarity_Index_", nationality, ".png"),
    plot = diss_map,
    width = 12,
    height = 8
  )
}
```

### Borough-Level Statistics for Dissimilarity Index

```{r}
borough_stats <- msoa_census_final %>%
  group_by(LAD11NM, Nationality) %>%
  summarize(
    mean_Diss_Index = mean(Diss_Index, na.rm = TRUE),
    min_Diss_Index = min(Diss_Index, na.rm = TRUE),
    max_Diss_Index = max(Diss_Index, na.rm = TRUE),
    .groups = "drop"
  )

print(borough_stats)
```

### Identify MSOAs and Boroughs with Most Changes in Dissimilarity Index

```{r}
# Calculate change in Diss_Index over time
change_stats <- msoa_census_final %>%
  filter(!is.na(Diss_Index), !is.na(Nationality)) %>%  # Remove rows with NA Diss_Index or Nationality
  group_by(MSOA11CD.x, LAD11NM, Nationality) %>%
  summarize(
    Diss_Index_Change = max(Diss_Index) - min(Diss_Index),  # No need for na.rm as NA rows are excluded
    .groups = "drop"
  )

# Top MSOAs where segregation DECREASED
top_msoas_decreased <- change_stats %>%
  filter(Diss_Index_Change < 0) %>%
  arrange(Diss_Index_Change) %>%
  group_by(Nationality) %>%
  slice_head(n = 5)

# View top MSOAs with decreased segregation
print(top_msoas_decreased)

# Top MSOAs where segregation INCREASED
top_msoas_increased <- change_stats %>%
  filter(Diss_Index_Change > 0) %>%
  arrange(desc(Diss_Index_Change)) %>%
  group_by(Nationality) %>%
  slice_head(n = 5)

# View top MSOAs with increased segregation
print(top_msoas_increased)

# Top Boroughs where segregation DECREASED
top_boroughs_decreased <- change_stats %>%
  filter(Diss_Index_Change < 0) %>%
  group_by(LAD11NM, Nationality) %>%
  summarize(
    Total_Change = sum(Diss_Index_Change),
    .groups = "drop"
  ) %>%
  arrange(Total_Change) %>%
  group_by(Nationality) %>%
  slice_head(n = 5)

# View top boroughs with decreased segregation
print(top_boroughs_decreased)

# Top Boroughs where segregation INCREASED
top_boroughs_increased <- change_stats %>%
  filter(Diss_Index_Change > 0) %>%
  group_by(LAD11NM, Nationality) %>%
  summarize(
    Total_Change = sum(Diss_Index_Change),
    .groups = "drop"
  ) %>%
  arrange(desc(Total_Change)) %>%
  group_by(Nationality) %>%
  slice_head(n = 5)

# View top boroughs with increased segregation
print(top_boroughs_increased)
```







