This document shows the code for the Exploratory Data Anlysis that will be done using the census demographic dataset.  

## Load necessary libraries

```{r}
# Install necessary libraries 
install.packages(c("dplyr", "ggplot2", "sf"))

# Load libraries
library(dplyr)
library(ggplot2)
library(sf)
library(readr)
```

## Load census data

```{r}
# Load the census demographic dataset
census_data <- read_csv("/Users/anamariasteinercorrea/thesis/census_data.csv")
msoa_geometry <- st_read("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/MSOA_2011_London_gen_MHW.shp")
```

## Adjust and correct data

```{r}
# View column names and types
str(census_data)

# Get a summary of key variables
summary(census_data)

# Check for missing values
sapply(census_data, function(x) sum(is.na(x)))
```

```{r}
# Remove rows with any NA values
census_data_clean <- na.omit(census_data)
```

```{r}
# Check if the dataset is an sf object
class(census_data_clean)

# Join to msoa shapefile
library(dplyr)
msoa_census <- msoa_geometry %>%
    left_join(census_data_clean, by = "MSOA11NM", "MSOA11CD")

colnames(msoa_census)

# Clean final columns
msoa_census_final <- msoa_census[, c("MSOA11CD.x", "MSOA11NM", "LAD11NM", "Year", "Nationality", "Proportion", "Diss_Index", "lagged_diss", "Theil_H", "Isolation", "geometry")]
```

```{r}
# Calculate statistics
min_value <- min(msoa_census_final$Diss_Index, na.rm = TRUE)
max_value <- max(msoa_census_final$Diss_Index, na.rm = TRUE)
median_value <- median(msoa_census_final$Diss_Index, na.rm = TRUE)
mean_value <- mean(msoa_census_final$Diss_Index, na.rm = TRUE)

# Print results
cat("Min:", min_value, "\n")
cat("Max:", max_value, "\n")
cat("Median:", median_value, "\n")
cat("Mean:", mean_value, "\n")
```
## EDA

### Visualice histograms and boxplots of distributions
```{r}
# Directory for histogram and boxplot outputs
dir.create("distribution_plots", showWarnings = FALSE)

# Histograms for all indices
for (var in c("Proportion", "Diss_Index", "Isolation", "Theil_H")) {
  hist_plot <- ggplot(msoa_census_final, aes_string(x = var)) +
    geom_histogram(bins = 30, fill = "blue", color = "black", alpha = 0.7) +
    facet_wrap(~ Nationality, scales = "free") +
    labs(title = paste("Histogram of", var, "by Nationality"),
         x = var, y = "Count") +
    theme_minimal()
  
  # Save histogram
  ggsave(filename = paste0("distribution_plots/Histogram_", var, ".png"),
         plot = hist_plot, width = 10, height = 6)
}

# Boxplots for all indices
for (var in c("Proportion", "Diss_Index", "Isolation", "Theil_H")) {
  box_plot <- ggplot(msoa_census_final, aes(x = Nationality, y = !!sym(var))) +
    geom_boxplot(fill = "lightblue", color = "black", alpha = 0.7) +
    labs(title = paste("Boxplot of", var, "by Nationality"),
         x = "Nationality", y = var) +
    theme_minimal()
  
  # Save boxplot
  ggsave(filename = paste0("distribution_plots/Boxplot_", var, ".png"),
         plot = box_plot, width = 10, height = 6)
}
```

### Visualice temporal line graphs of all indices
```{r}
library(ggplot2)
library(dplyr)

# Ensure Year is numeric
msoa_census_final$Year <- as.numeric(msoa_census_final$Year)

# Aggregate data by Year and Nationality
aggregated_data <- msoa_census_final %>%
  group_by(Year, Nationality) %>%
  summarize(
    Proportion = mean(Proportion, na.rm = TRUE),
    Diss_Index = mean(Diss_Index, na.rm = TRUE),
    Lagged_Diss = mean(lagged_diss, na.rm = TRUE),
    Theil_H = mean(Theil_H, na.rm = TRUE),
    Isolation = mean(Isolation, na.rm = TRUE),
    .groups = "drop"
  )

# Create directory for nationality-wise plots
dir.create("line_graphs_nationality", showWarnings = FALSE)

# Loop through each nationality and create line plots
nationalities <- unique(aggregated_data$Nationality)

for (nationality in nationalities) {
  # Filter data for the current nationality
  nat_data <- aggregated_data %>%
    filter(Nationality == nationality)

  # Generate the line plot
  plot <- ggplot(nat_data) +
    geom_line(aes(x = Year, y = Proportion, color = "Proportion"), size = 1.2) +
    geom_line(aes(x = Year, y = Diss_Index, color = "Dissimilarity Index"), size = 1.2) +
    geom_line(aes(x = Year, y = Theil_H, color = "Theil's H"), size = 1.2) +
    geom_line(aes(x = Year, y = Isolation, color = "Isolation Index"), size = 1.2) +
    labs(
      title = paste("Temporal Trends for", nationality),
      x = "Year", y = "Value", color = "Index"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      legend.position = "top"
    )

  # Save the plot
  ggsave(
    filename = paste0("line_graphs_nationality/Trends_", nationality, ".png"),
    plot = plot,
    width = 10,
    height = 6
  )
}
```

### Visualice and compare temporal line graph between all nationalities
```{r}
# Create line graph for Dissimilarity Index comparing all nationalities
dissimilarity_plot <- ggplot(aggregated_data, aes(x = Year, y = Diss_Index, color = Nationality, group = Nationality)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(title = "Dissimilarity Index Over Time (All Nationalities)",
       x = "Year", y = "Dissimilarity Index", color = "Nationality") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

# Save the plot
dir.create("line_graphs_comparison", showWarnings = FALSE)
ggsave(filename = "line_graphs_comparison/Dissimilarity_Index_All_Nationalities.png",
       plot = dissimilarity_plot, width = 10, height = 6)
```

### Compare heatmaps of diss index between nationalities
```{r}
# Directory for spatial heatmaps of diss index
dir.create("maps", showWarnings = FALSE)

# Years to map and Nationalities
years_to_map <- c(2001, 2011, 2021)
nationalities <- unique(msoa_census_final$Nationality)

# Generate heatmaps
for (nationality in nationalities) {
  for (year in years_to_map) {
    map_data <- msoa_census_final %>%
      filter(Nationality == nationality, Year == year) %>%
      st_as_sf()
    
    map_plot <- ggplot(map_data) +
      geom_sf(aes(fill = Diss_Index), color = NA) +
      scale_fill_viridis_c(option = "C") +
      labs(title = paste("Dissimilarity Index for", nationality, "-", year),
           fill = "Dissimilarity Index") +
      theme_minimal()
    
    # Save map
    ggsave(filename = paste0("maps/Dissimilarity_Map_", nationality, "_", year, ".png"),
           plot = map_plot, width = 10, height = 6)
  }
}
```

### Calculate spatial autocorrelation of diss index between nationalities
```{r}
library(sf)
library(dplyr)

# Directory for outputs
dir.create("moran_outputs", showWarnings = FALSE)

# Loop through each nationality
for (nationality in nationalities) {
  # Filter data for the current nationality
  nat_data <- msoa_census_final %>%
    filter(Nationality == nationality) %>%
    st_as_sf()  # Ensure it's an sf object
  
  # Create spatial weights matrix (binary adjacency based on geometry)
  coords <- st_coordinates(st_centroid(nat_data))  # Centroids for spatial locations
  dist_matrix <- as.matrix(dist(coords))           # Distance matrix
  
  # Create binary adjacency matrix: 1 if neighbors, 0 otherwise (threshold: small distance)
  threshold <- quantile(dist_matrix, 0.1)  # Use the 10th percentile as neighbor threshold
  weights <- ifelse(dist_matrix > 0 & dist_matrix <= threshold, 1, 0)
  
  # Moran's I calculation
  x <- nat_data$Diss_Index                       # Values
  x_mean <- mean(x, na.rm = TRUE)                # Mean of Diss_Index
  n <- length(x)                                 # Number of observations
  W <- sum(weights)                              # Total weight sum
  
  numerator <- sum(weights * outer(x - x_mean, x - x_mean))  # Weighted cross-product
  denominator <- sum((x - x_mean)^2)                         # Variance
  
  moran_I <- (n / W) * (numerator / denominator)             # Moran's I statistic
  
  # Save Moran's I results
  writeLines(paste("Moran's I for", nationality, "\n",
                   "Moran's I statistic:", round(moran_I, 4), "\n"),
             con = paste0("moran_outputs/Moran_", nationality, ".txt"))
  
  # Moran Scatterplot
  lagged_values <- (weights %*% x) / rowSums(weights)  # Spatially lagged values
  
  moran_plot <- ggplot(data = data.frame(Diss_Index = x, Lagged = lagged_values)) +
    geom_point(aes(x = Diss_Index, y = Lagged), alpha = 0.5, color = "blue") +
    labs(title = paste("Moran's Scatterplot for", nationality),
         x = "Dissimilarity Index",
         y = "Spatially Lagged Dissimilarity Index") +
    theme_minimal()
  
  # Save Moran scatterplot
  ggsave(filename = paste0("moran_outputs/Moran_Scatter_", nationality, ".png"),
         plot = moran_plot, width = 8, height = 6)
}
```

```{r}
set.seed(123)  # Ensure reproducibility

# Number of random permutations
n_simulations <- 999  

# Run for each nationality
for (nationality in nationalities) {
  # Filter data for the current nationality
  nat_data <- msoa_census_final %>%
    filter(Nationality == nationality) %>%
    st_as_sf()

  # Create spatial weights
  coords <- st_coordinates(st_centroid(nat_data))
  dist_matrix <- as.matrix(dist(coords))
  threshold <- quantile(dist_matrix, 0.1)
  weights <- ifelse(dist_matrix > 0 & dist_matrix <= threshold, 1, 0)

  # Original Moran's I calculation
  x <- nat_data$Diss_Index
  x_mean <- mean(x, na.rm = TRUE)
  n <- length(x)
  W <- sum(weights)
  numerator <- sum(weights * outer(x - x_mean, x - x_mean))
  denominator <- sum((x - x_mean)^2)
  observed_moran_I <- (n / W) * (numerator / denominator)

  # Permutation test
  simulated_moran_I <- numeric(n_simulations)

  for (i in 1:n_simulations) {
    shuffled_x <- sample(x)  # Shuffle Dissimilarity Index values
    numerator_perm <- sum(weights * outer(shuffled_x - mean(shuffled_x), shuffled_x - mean(shuffled_x)))
    simulated_moran_I[i] <- (n / W) * (numerator_perm / denominator)
  }

  # Calculate p-value
  p_value <- mean(simulated_moran_I >= observed_moran_I)

  # Save Moran's I results with p-value
  writeLines(paste("Moran's I for", nationality, "\n",
                   "Moran's I statistic:", round(observed_moran_I, 4), "\n",
                   "P-value:", round(p_value, 4), "\n"),
             con = paste0("moran_outputs/Moran_", nationality, ".txt"))

  # Print results to console
  cat("Nationality:", nationality, "\n",
      "Observed Moran's I:", round(observed_moran_I, 4), "\n",
      "P-value:", round(p_value, 4), "\n\n")
}
```
```{r}
Moran's I	Meaning
> 0	Positive spatial autocorrelation (clustering)
< 0	Negative spatial autocorrelation (dispersion)
≈ 0	No spatial autocorrelation (random)
P-Value	Interpretation
< 0.05	Significant spatial autocorrelation
> 0.05	No significant spatial autocorrelation
```

### Visualice maps of diss index between nationalities and through time
```{r}
# Create directory for spatial maps
dir.create("spatial_maps", showWarnings = FALSE)

# Plot map for Dissimilarity Index for all MSOAs
diss_map <- ggplot(msoa_census_final) +
  geom_sf(aes(fill = Diss_Index, geometry = geometry), color = NA) +
  scale_fill_viridis_c(option = "C", name = "Dissimilarity Index") +
  labs(title = "Dissimilarity Index Across London (All Nationalities)",
       caption = "Source: Census Data") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        legend.position = "right")

# Save the plot
ggsave(filename = "spatial_maps/Dissimilarity_Index_London.png",
       plot = diss_map, width = 12, height = 8)
```

### Borough-Level Statistics for Dissimilarity Index

```{r}
borough_stats <- msoa_census_final %>%
  group_by(LAD11NM, Nationality) %>%
  summarize(
    mean_Diss_Index = mean(Diss_Index, na.rm = TRUE),
    min_Diss_Index = min(Diss_Index, na.rm = TRUE),
    max_Diss_Index = max(Diss_Index, na.rm = TRUE),
    .groups = "drop"
  )

print(borough_stats)
```

### Identify MSOAs and Boroughs with Most Changes in Dissimilarity Index

```{r}
# Calculate change in Diss_Index over time
change_stats <- msoa_census_final %>%
  group_by(MSOA11CD.x, LAD11NM, Nationality) %>%
  summarize(
    Diss_Index_Change = max(Diss_Index, na.rm = TRUE) - min(Diss_Index, na.rm = TRUE),
    .groups = "drop"
  )

# Top MSOAs with most changes in Dissimilarity Index
top_msoas <- change_stats %>%
  arrange(desc(Diss_Index_Change)) %>%
  group_by(Nationality) %>%
  slice_head(n = 10)  # Top 10 MSOAs for each nationality

# View top MSOAs
print(top_msoas)

# Top Boroughs with most changes in Dissimilarity Index
top_boroughs <- change_stats %>%
  group_by(LAD11NM, Nationality) %>%
  summarize(
    Total_Change = sum(Diss_Index_Change, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(Total_Change)) %>%
  group_by(Nationality) %>%
  slice_head(n = 10)  # Top 10 boroughs for each nationality

# View top boroughs
print(top_boroughs)
```







