Stepwise Analysis:
Helps evaluate the incremental effect of adding the policy and interaction terms.
Control for Baseline:
Ensures the policy effect is isolated after accounting for other key factors.
Interpretability:
Breaks down complex interactions into manageable components.

Baseline Model:
Captures the effects of housing variables and London-wide average Y
Policy Model:
Adds the policy dummy to test its direct effect on Y after controlling for baseline variables.
Interaction Model:
Adds interaction terms to explore whether the policy modifies the relationships between housing variables and 
Y.

```{r}
# Required Libraries
library(tidyverse)
library(car) # For VIF and diagnostics
library(stargazer) # For traditional regression tables
```

```{r}
housing_data <- read_csv("/Users/anamariasteinercorrea/thesis/housing_data.csv")
neighborhood_data <- read_csv("/Users/anamariasteinercorrea/thesis/neighborhood_data.csv")
msoa_geometry <- st_read("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/MSOA_2011_London_gen_MHW.shp")
```

```{r}
# Perform the join of housing + neighborhood
hn_data <- housing_data %>%
  left_join(neighborhood_data, by = c("MSOA11NM", "MSOA11CD", "Year"))
```

```{r}
# Join with spatial data
hn_final <- hn_data %>%
  left_join(msoa %>% distinct(MSOA11CD, MSOA11NM, .keep_all = TRUE), 
            by = c("MSOA11CD", "MSOA11NM"))

# Convert to sf object
dummies_sf <- st_as_sf(hn_final)

# Clean the dataset
final_dummies <- dummies_sf %>%
  # Select only the necessary columns and rename them
  select(
    MSOA11CD,
    MSOA11NM,
    LAD11NM,
    Year,
    Rent,
    budget_proportion,
    average_owned,
    Houseprice,
    Occupation,
    geometry = geometry.y,
    MSOA_crime,
    MSOA_percent_green,
    PTAL
  ) 
```

```{r}
# Filter rows for "London" and get the unique values for transformation
city_of_london <- final_dummies[final_dummies$MSOA11NM == "London", ]

# Rename columns for clarity
city_of_london <- city_of_london %>%
  rename(
    London_occupation = Occupation,
    London_rent = Rent,
    London_average_owned = average_owned,
    London_houseprice = Houseprice
  )

# Ensure there is only one unique value per year for the London columns
city_of_london <- city_of_london %>%
  group_by(Year) %>%
  slice(1) %>%
  ungroup()

# Select only necessary columns from "City of London" and prepare to merge
city_of_london <- city_of_london %>%
  select(Year, London_occupation, London_rent, London_average_owned, London_houseprice)

# Drop geometry from city_of_london
city_of_london <- city_of_london %>%
  st_drop_geometry()

# Perform the join based on "Year"
merged_dummies <- final_dummies %>%
  left_join(city_of_london, by = "Year")

# Drop rows corresponding to "London"
merged_dummies <- merged_dummies %>%
  filter(MSOA11NM != "London")
```

## Dummie policy data
```{r}
# Adjusted policy definitions for 2001–2021 with pre and post = 0, during = 1
policy <- list(
  "Right_to_Buy" = list(
    boroughs = c("Barking", "Havering", "Redbridge", "Newham", "Haringey"),
    year_ranges = list(pre_post = c(2001:2003, 2010:2021), during = 2004:2009)
  ),
  "Welfare_Reform_Act" = list(
    boroughs = c("Newham", "Tower Hamlets", "Barking and Dagenham", "Hackney", "Enfield"),
    year_ranges = list(pre_post = c(2001:2011, 2017:2021), during = 2012:2016)
  ),
  "Elephant_and_Castle_Redevelopment" = list(
    boroughs = c("Southwark"),
    year_ranges = list(pre_post = c(2001:2003, 2022), during = 2004:2021)
  ),
  "Affordable_Housing_Program" = list(
    boroughs = c("Newham", "Tower Hamlets", "Croydon", "Barking"),
    year_ranges = list(pre_post = c(2001:2015, 2022), during = 2016:2021)
  ),
  "Housing_Act_2004" = list(
    boroughs = c("Tower Hamlets", "Hackney", "Newham", "Brent", "Ealing"),
    year_ranges = list(pre_post = c(2001:2003, 2012:2021), during = 2004:2011)
  )
)

# Add dummy variables for each policy
for (policy_name in names(policy)) {
  policy_info <- policy[[policy_name]]
  
  # Add a new column for each policy
  merged_dummies[[policy_name]] <- with(merged_dummies,
    ifelse(
      LAD11NM %in% policy_info$boroughs & Year %in% policy_info$year_ranges$during, 
      1, 
      0
    )
  )
}
```

```{r}
library(dplyr)
library(sf)         # For st_drop_geometry()
library(ggcorrplot) # For correlation plot

# Define columns of interest, including public policy variables
correlation_columns <- c(
  "Rent", "budget_proportion", "average_owned", "Houseprice", "Occupation", 
  "London_occupation", "London_rent", "London_average_owned", "London_houseprice",
  "MSOA_crime", "MSOA_percent_green", "PTAL",
  "Right_to_Buy", "Welfare_Reform_Act", "Elephant_and_Castle_Redevelopment",
  "Affordable_Housing_Program", "Housing_Act_2004"
)

# Step 1: Drop geometry and filter numeric columns
numeric_data <- merged_dummies %>%
  st_drop_geometry() %>%
  select(all_of(correlation_columns)) %>%
  select(where(is.numeric))

# Step 2: Compute and save the correlation matrix
correlation_dummies <- cor(numeric_data, use = "complete.obs")

# Step 3: Plot the correlation matrix
correlation_plot <- ggcorrplot(
  correlation_dummies,
  hc.order = TRUE,
  type = "lower",
  lab = TRUE,
  title = "Correlation Matrix for Housing, Neighborhood, and Public Policy Variables"
)

# Step 4: Save the plot (optional)
output_file <- "correlation_matrix.png"
ggsave(output_file, plot = correlation_plot, width = 10, height = 8)
```

```{r}

library(dplyr)
library(sf)
library(car)
library(stargazer)
library(sjPlot)

# Create directory for regression outputs
output_dir <- "housing_regression_outputs"
if (!dir.exists(output_dir)) dir.create(output_dir)

# Define regression details for each housing characteristic
housing_characteristics <- list(
  list(
    name = "average_owned",
    y_var = "average_owned",
    match_var = "London_average_owned",
    policy_dummy = "Affordable_Housing_Program",
    x_vars = c("Rent", "budget_proportion", "Houseprice", "Occupation")
  ),
  list(
    name = "rent",
    y_var = "Rent",
    match_var = "London_rent",
    policy_dummy = "Right_to_Buy",
    x_vars = c("budget_proportion", "average_owned", "Houseprice", "Occupation")
  ),
  list(
    name = "houseprice",
    y_var = "Houseprice",
    match_var = "London_houseprice",
    policy_dummy = "Elephant_and_Castle_Redevelopment",
    x_vars = c("Rent", "budget_proportion", "average_owned", "Occupation")
  ),
  list(
    name = "occupation",
    y_var = "Occupation",
    match_var = "London_occupation",
    policy_dummy = "Welfare_Reform_Act",
    x_vars = c("Rent", "budget_proportion", "average_owned", "Houseprice")
  ),
  list(
    name = "budget_proportion",
    y_var = "budget_proportion",
    match_var = "London_average_owned",
    policy_dummy = "Housing_Act_2004",
    x_vars = c("Rent", "average_owned", "Houseprice", "Occupation")
  )
)

# Loop through each housing characteristic and run the regressions
for (housing in housing_characteristics) {
  name <- housing$name
  y_var <- housing$y_var
  match_var <- housing$match_var
  policy_dummy <- housing$policy_dummy
  x_vars <- c(housing$x_vars, match_var)
  
  # Drop geometry and filter missing data
  data_clean <- merged_dummies %>%
    st_drop_geometry() %>%
    filter(complete.cases(select(., all_of(c(y_var, x_vars, policy_dummy)))))
  
  # Progression 1: Base Model
  base_formula <- as.formula(paste(y_var, "~", paste(x_vars, collapse = " + ")))
  base_model <- lm(base_formula, data = data_clean)
  
  # Progression 2: Policy Dummy Model
  dummy_formula <- as.formula(paste(y_var, "~", paste(x_vars, collapse = " + "), "+", policy_dummy))
  dummy_model <- lm(dummy_formula, data = data_clean)
  
  # Progression 3: Interaction Model
  interaction_terms <- paste(paste0(x_vars, "*", policy_dummy), collapse = " + ")
  interaction_formula <- as.formula(paste(y_var, "~", paste(x_vars, collapse = " + "), "+", policy_dummy, "+", interaction_terms))
  interaction_model <- lm(interaction_formula, data = data_clean)
  
  # Check for aliased coefficients
  aliased <- alias(interaction_model)
  if (length(aliased$Complete) > 0) {
    print(paste("Aliased coefficients detected for", name, ". Skipping VIF calculation."))
    print(aliased$Complete)
  } else {
    # Multicollinearity check for interaction model
    vif_values <- vif(interaction_model)
    print(paste("VIF for", name, ":"))
    print(vif_values)
  }
  
  # Save diagnostic plots
  diagnostics_file <- file.path(output_dir, paste0(name, "_Diagnostics.pdf"))
  pdf(diagnostics_file)
  par(mfrow = c(2, 2))
  plot(base_model)
  plot(dummy_model)
  plot(interaction_model)
  dev.off()
  
  # Save regression tables
  regression_summary_file <- file.path(output_dir, paste0(name, "_RegressionSummary.html"))
  stargazer(
    base_model, dummy_model, interaction_model,
    type = "html",
    out = regression_summary_file,
    title = paste("Regression Results for", name),
    digits = 3,
    star.cutoffs = c(0.05, 0.01, 0.001)
  )
  
  # Save styled regression table
  styled_summary_file <- file.path(output_dir, paste0(name, "_StyledRegressionSummary.html"))
  tab_model(
    base_model, dummy_model, interaction_model,
    file = styled_summary_file,
    show.ci = TRUE, show.std = TRUE,
    dv.labels = c("Base Model", "Policy Dummy Model", "Interaction Model"),
    title = paste("Styled Regression Results for", name)
  )
}
```

