## Load libraries

```{r}
# Install all necessary packages
install.packages(c("tidyverse", "sf", "sp", "spgwr", "rgdal", "raster", 
                   "tmap", "ggplot2", "gridExtra", "cowplot", "xts", "zoo"))

# Load all packages
library(tidyverse)
library(sf)
library(sp)
library(spgwr)
library(raster)
library(tmap)
library(ggplot2)
library(gridExtra)
library(cowplot)
library(xts)
library(zoo)
```

```{r}
# Read the datasets
setwd("/Users/anamariasteinercorrea/thesis")
census_data <- read_csv("/Users/anamariasteinercorrea/thesis/census_data.csv")
housing_data <- read_csv("/Users/anamariasteinercorrea/thesis/housing_data.csv")
neighborhood_data <- read_csv("/Users/anamariasteinercorrea/thesis/neighborhood_data.csv")
msoa_geometry <- st_read("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/MSOA_2011_London_gen_MHW.shp")
```

## Clean and join datasets

```{r}
# Perform the join of housing + neighborhood
hn_data <- housing_data %>%
  left_join(neighborhood_data, by = c("MSOA11NM", "MSOA11CD", "Year"))
```

```{r}
# Perform the join of housing + neighborhood + segregation
segregation_analysis <- hn_data %>%
  left_join(census_data, by = c("MSOA11NM", "MSOA11CD", "Year"))
```

```{r}
# Join with spatial data
segregation_final <- segregation_analysis %>%
  left_join(msoa %>% distinct(MSOA11CD, MSOA11NM, .keep_all = TRUE), 
            by = c("MSOA11CD", "MSOA11NM"))

# Convert to sf object
segregation_sf <- st_as_sf(segregation_final)

# Clean the dataset
final_segregation <- segregation_sf %>%
  # Select only the necessary columns and rename them
  select(
    MSOA11CD,
    MSOA11NM,
    Year,
    Rent,
    budget_proportion,
    average_owned,
    average_rented,
    average_social,
    Houseprice,
    Occupation,
    geometry = geometry.y,
    MSOA_crime,
    MSOA_percent_green,
    PTAL,
    Nationality,
    Dissimilarity_Index,
    lagged_diss = lagged_diss.y,
    H,
    Isolation_Index,
  ) 

filtered_segregation<- final_segregation %>%
  filter(MSOA11NM != "London")
```

## Question 1: How do housing characteristics affect segregation?

## Scatterplots
To visualize the instantaneous relationships between each housing variable and the Dissimilarity Index for each nationality.

```{r}
# Define nationalities and housing variables
nationalities <- c("Bangladesh", "Caribbean", "Poland", "Romania", "SEAfrica", "SouthAmerica")
housing_variables <- c("Rent", "budget_proportion", "average_owned", "Houseprice", "Occupation", "lagged_diss")

# Create directories for scatterplots
dir.create("scatter_plots", showWarnings = FALSE)

install.packages("ggplot2")
library(ggplot2)

# Loop through nationalities and housing variables to create scatterplots
for (nationality in nationalities) {
  # Filter data for the current nationality
  nationality_data <- filtered_segregation %>% filter(Nationality == nationality)
  
  for (housing_var in housing_variables) {
    # Create scatterplot
    scatter_plot <- ggplot(nationality_data, aes_string(x = housing_var, y = "Dissimilarity_Index")) +
      geom_point(color = "blue", alpha = 0.5) +
      geom_smooth(method = "lm", color = "red", se = FALSE) +
      labs(
        title = paste("Scatter Plot:", housing_var, "vs Dissimilarity Index -", nationality),
        x = housing_var,
        y = "Dissimilarity Index"
      ) +
      theme_minimal()
    
    # Save scatterplot
    ggsave(
      filename = paste0("scatter_plots/Scatter_", nationality, "_", housing_var, ".png"), 
      plot = scatter_plot, 
      width = 8, 
      height = 6
    )
  }
}
```

## Correlation matrix
To summarize the strength and direction of pairwise relationships between all housing variables and the Dissimilarity Index for each nationality.
```{r}
# Load required libraries
library(dplyr)

# Define nationalities
nationalities <- c("Bangladesh", "Caribbean", "Poland", "Romania", "SEAfrica", "SouthAmerica")

# Directory for correlation matrices
dir.create("correlation_matrices", showWarnings = FALSE)

install.packages("ggcorrplot")
library(ggcorrplot)

# Loop through nationalities to generate correlation matrices
for (nationality in nationalities) {
  # Filter data for the current nationality
  nationality_data <- filtered_segregation %>% filter(Nationality == nationality)
  
  # Drop the geometry column using st_drop_geometry()
  relevant_data <- nationality_data %>%
    st_drop_geometry() %>%  # Drop geometry column
    select(c("Rent", "budget_proportion", "average_owned", "Houseprice", "Occupation", "Dissimilarity_Index", "lagged_diss")) %>%
    mutate(across(where(is.character), as.numeric))  # Convert character columns to numeric
  
  # Ensure all columns are numeric
  numeric_data <- relevant_data %>%
    select(where(is.numeric))  # Keep only numeric columns
  
  # Check if there are enough numeric columns and rows
  if (ncol(numeric_data) > 1 && nrow(numeric_data) > 1) {
    # Compute correlation matrix
    cor_matrix <- cor(numeric_data, use = "complete.obs")
    
    # Visualize correlation matrix as a heatmap
    cor_plot <- ggcorrplot(cor_matrix, method = "circle", type = "lower", lab = TRUE, 
                           title = paste("Correlation Matrix -", nationality),
                           colors = c("blue", "white", "red"))
    
    # Save the correlation matrix plot
    ggsave(filename = paste0("correlation_matrices/Correlation_", nationality, ".png"), 
           plot = cor_plot, width = 8, height = 6)
  } else {
    message(paste("Skipping nationality:", nationality, "- insufficient numeric data for correlation."))
  }
}
```

## Cross correlation function analysis
CCF for all six nationalities and all five x variables (housing characteristics). This will result in 30 CCF plots (6 nationalities × 5 variables).
```{r}
# Load required libraries
library(tidyverse)

# Define nationalities and housing variables
nationalities <- c("Bangladesh", "Caribbean", "Poland", "Romania", "SEAfrica", "SouthAmerica")
housing_variables <- c("Rent", "budget_proportion", "average_owned", "Houseprice", "Occupation", "lagged_diss")

# Create a directory to save CCF plots if it doesn't exist
dir.create("ccf_plots", showWarnings = FALSE)

# Loop over nationalities and housing variables
for (nationality in nationalities) {
  # Filter data for the current nationality
  nationality_data <- filtered_segregation %>% filter(Nationality == nationality)
  
  # Ensure data is sorted by Year
  nationality_data <- nationality_data %>% mutate(Year = as.numeric(Year)) %>% arrange(Year)
  
  # Extract the segregation index as the time series (dependent variable)
  segregation_ts <- ts(nationality_data$Dissimilarity_Index, start = min(nationality_data$Year), frequency = 1)
  
  for (housing_var in housing_variables) {
    # Extract the housing variable as the time series (independent variable)
    housing_ts <- ts(nationality_data[[housing_var]], start = min(nationality_data$Year), frequency = 1)
    
    # Check for and handle missing values
    if (anyNA(segregation_ts) || anyNA(housing_ts)) {
      # Filter out rows with NA values in either series
      valid_data <- nationality_data %>% 
        filter(!is.na(Dissimilarity_Index) & !is.na(.data[[housing_var]]))
      
      # Recreate the time series with non-NA values
      segregation_ts <- ts(valid_data$Dissimilarity_Index, start = min(valid_data$Year), frequency = 1)
      housing_ts <- ts(valid_data[[housing_var]], start = min(valid_data$Year), frequency = 1)
    }
    
    # Save the CCF plot as an image
    png(filename = paste0("ccf_plots/CCF_", nationality, "_", housing_var, ".png"), width = 800, height = 600)
    ccf(housing_ts, segregation_ts, lag.max = 5, 
        main = paste("CCF:", housing_var, "vs Dissimilarity Index -", nationality))
    dev.off()
  }
}
```

# Base regressions (1-5 housing characteristics)
Run separate regressions for each housing variable to assess their individual contributions to the segregation index.
##No. 6 Lagged settlement model
Incorporate the lagged settlement variable alongside all housing variables to capture temporal dependencies.
##No. 7 Neighborhood data
Add neighborhood-level variables to create a comprehensive model.
```{r}
# Load required libraries
library(tidyverse)
library(car) 
install.packages("stargazer")
library(stargazer)
install.packages("sjPlot")
library(sjPlot)    # For styled regression tables

# Define variables
housing_variables <- c("Rent", "budget_proportion", "average_owned", "Houseprice", "Occupation")
neighborhood_variables <- c("MSOA_crime", "MSOA_percent_green", "PTAL")
nationalities <- c("Bangladesh", "Caribbean", "Poland", "Romania", "SEAfrica", "SouthAmerica")

# Create directory for regression outputs
dir.create("regression_outputs", showWarnings = FALSE)

# Loop over nationalities
for (nationality in nationalities) {
  # Filter data for the current nationality
  nationality_data <- filtered_segregation %>% filter(Nationality == nationality)
  
  # Drop geometry column
  numeric_data <- st_drop_geometry(nationality_data)
  
  # Handle missing values
  numeric_data <- numeric_data %>% filter(complete.cases(select(., all_of(c(housing_variables, neighborhood_variables, "lagged_diss")))))
  
  # Scale continuous variables
  scaled_data <- numeric_data %>%
    mutate(across(all_of(c(housing_variables, neighborhood_variables, "lagged_diss")), scale, .names = "scaled_{.col}"))
  
  # Define dependent variable
  y_variable <- "Dissimilarity_Index"
  
  # Initialize list to store models
  models <- list()
  
  # 1. Regressions for individual housing variables
  for (housing_var in housing_variables) {
    formula <- as.formula(paste(y_variable, "~", paste0("scaled_", housing_var)))
    model <- lm(formula, data = scaled_data)
    models[[paste0("Housing_", housing_var)]] <- model
  }
  
  # 2. Model with lagged settlement
  lagged_formula <- as.formula(paste(y_variable, "~", 
                                     paste(paste0("scaled_", housing_variables), collapse = " + "), 
                                     "+ scaled_lagged_diss"))
  models[["WithLaggedSettlement"]] <- lm(lagged_formula, data = scaled_data)
  
  # 3. Full model with neighborhood variables
  full_formula <- as.formula(paste(y_variable, "~", 
                                   paste(paste0("scaled_", housing_variables), collapse = " + "), 
                                   "+ scaled_lagged_diss +", 
                                   paste(paste0("scaled_", neighborhood_variables), collapse = " + ")))
  models[["FullModel"]] <- lm(full_formula, data = scaled_data)
  
  # Check for multicollinearity in the full model
  vif_values <- vif(models[["FullModel"]])
  print(vif_values)  # Check if VIF > 5–10
  
  # Save diagnostics plots
  pdf(file = paste0("regression_outputs/Diagnostics_", nationality, "_FullModel.pdf"))
  par(mfrow = c(2, 2))
  plot(models[["FullModel"]])  # Residuals, QQ, leverage
  dev.off()
  
  # Generate regression tables
  try({
    stargazer(models, 
              type = "html", 
              out = paste0("regression_outputs/RegressionSummary_", nationality, ".html"), 
              title = paste("Regression Results for", nationality), 
              digits = 3, 
              star.cutoffs = c(0.05, 0.01, 0.001))
  }, silent = TRUE)
  
  # Styled regression table
  try({
    tab_model(models, 
              file = paste0("regression_outputs/RegressionSummary_", nationality, "_Styled.html"), 
              show.ci = TRUE, show.std = TRUE, dv.labels = paste("Dissimilarity Index -", nationality))
  }, silent = TRUE)
}
```

# GWR with best regression model
Explores spatial variations in these relationships, revealing local heterogeneity.
```{r}
# Required Libraries
library(sf)
install.packages("spgwr")
library(spgwr)
library(tidyverse)

# Define variables
nationalities <- c("Bangladesh", "Caribbean", "Poland", "Romania", "SEAfrica", "SouthAmerica")
best_formula <- paste("Dissimilarity_Index ~ Rent + budget_proportion + average_owned + Houseprice + lagged_diss + Occupation + MSOA_crime + MSOA_percent_green + PTAL")

# Ensure directory for GWR outputs
dir.create("gwr_outputs", showWarnings = FALSE)

# Deduplicate geometries (prepare unique geometries for MSOAs)
unique_geometries <- filtered_segregation %>%
  select(MSOA11CD, geometry) %>%
  distinct()  # Retain only unique geometries

# Loop over nationalities
for (nationality in nationalities) {
  # Filter data for the current nationality
  nat_data <- filtered_segregation %>%
    filter(Nationality == nationality) %>%
    filter(!is.na(geometry)) %>%
    group_by(MSOA11CD) %>%
    summarize(across(where(is.numeric), mean, na.rm = TRUE), .groups = "drop") %>%
    left_join(unique_geometries, by = "MSOA11CD") %>%  # Reattach unique geometries
    st_as_sf() %>%
    st_make_valid()  # Ensure valid geometries
  
  # Convert to Spatial* object for spgwr
  nat_spatial <- as(nat_data, "Spatial")

  # Bandwidth selection (adaptive GWR)
  bandwidth <- gwr.sel(
    as.formula(best_formula),
    data = nat_spatial,
    coords = coordinates(nat_spatial),
    gweight = gwr.Gauss
  )

  # Run GWR
  gwr_model <- gwr(
    as.formula(best_formula),
    data = nat_spatial,
    coords = coordinates(nat_spatial),
    bandwidth = bandwidth,
    gweight = gwr.Gauss,
    fit.points = TRUE
  )

  # Extract GWR results
  gwr_results <- as.data.frame(gwr_model$SDF)

  # Save results back to spatial data
  nat_data <- cbind(nat_data, gwr_results)

  # Save GWR output to a shapefile
  st_write(nat_data, paste0("gwr_outputs/GWR_", nationality, ".shp"), delete_dsn = TRUE)

  # Visualize GWR coefficient maps for selected predictors
  for (variable in c("Rent", "budget_proportion", "average_owned", "Houseprice", "lagged_diss", "Occupation", "MSOA_crime", "MSOA_percent_green", "PTAL")) {
    ggplot(nat_data) +
      geom_sf(aes_string(fill = variable), color = NA) +
      scale_fill_viridis_c() +
      labs(
        title = paste("GWR Coefficient Map for", variable, "(", nationality, ")"),
        fill = "Coefficient"
      ) +
      theme_minimal() +
      ggsave(paste0("gwr_outputs/GWR_Coefficient_", nationality, "_", variable, ".png"), width = 8, height = 6)
  }
}
```

# Diagnostico
```{r}
# Calculate RMSE for GWR
observed <- gwr_model$SDF$Dissimilarity_Index # Replace with your dependent variable's name
predicted <- gwr_model$SDF$gwr.pred           # GWR-predicted values
rmse <- sqrt(mean((observed - predicted)^2, na.rm = TRUE))

# Print and save RMSE
cat("RMSE for GWR (", nationality, "):", rmse, "\n")
write.table(data.frame(Nationality = nationality, RMSE = rmse),
            file = paste0("gwr_outputs/RMSE_", nationality, ".txt"), row.names = FALSE, append = TRUE)

# Additional diagnostics for each GWR model
for (nationality in nationalities) {
  # Filter data for the current nationality
  nat_data <- spatial_data %>% filter(Nationality == nationality)
  
  # Ensure spatial features are valid
  nat_data <- nat_data %>% st_make_valid()
  
  # Convert to Spatial* object (required by spgwr)
  nat_spatial <- as(nat_data, "Spatial")
  
  # Bandwidth selection (adaptive GWR)
  bandwidth <- gwr.sel(as.formula(best_formula), data = nat_spatial, coords = coordinates(nat_spatial), gweight = gwr.Gauss)
  
  # Run GWR
  gwr_model <- gwr(as.formula(best_formula), data = nat_spatial, coords = coordinates(nat_spatial), bandwidth = bandwidth, gweight = gwr.Gauss, fit.points = TRUE)
  
  # Extract GWR results
  gwr_results <- as.data.frame(gwr_model$SDF)
  local_r2 <- gwr_results$localR2
  residuals <- gwr_results$residuals
  
  # Save results back to spatial data
  nat_data <- cbind(nat_data, gwr_results)
  
  # Save GWR output to a shapefile
  st_write(nat_data, paste0("gwr_outputs/GWR_", nationality, ".shp"), delete_dsn = TRUE)
  
  # Plot coefficient maps for selected predictors
  for (variable in c("HousingIndex", "RentPrice", "Tenancy", "LaggedSettlement")) {
    ggplot(nat_data) +
      geom_sf(aes_string(fill = variable), color = NA) +
      scale_fill_viridis_c() +
      labs(title = paste("GWR Coefficient Map for", variable, "(", nationality, ")"), fill = "Coefficient") +
      theme_minimal() +
      ggsave(paste0("gwr_outputs/GWR_Coefficient_", nationality, "_", variable, ".png"), width = 8, height = 6)
  }
  
  # Plot local R-squared
  ggplot(nat_data) +
    geom_sf(aes(fill = local_r2), color = NA) +
    scale_fill_viridis_c() +
    labs(title = paste("Local R-squared Map (", nationality, ")"), fill = "R-squared") +
    theme_minimal() +
    ggsave(paste0("gwr_outputs/GWR_LocalR2_", nationality, ".png"), width = 8, height = 6)
  
  # Residual diagnostics
  ggplot(nat_data) +
    geom_sf(aes(fill = residuals), color = NA) +
    scale_fill_viridis_c() +
    labs(title = paste("Residual Map (", nationality, ")"), fill = "Residuals") +
    theme_minimal() +
    ggsave(paste0("gwr_outputs/GWR_Residuals_", nationality, ".png"), width = 8, height = 6)
  
  # Spatial autocorrelation of residuals (Moran's I)
  moran <- spdep::moran.test(residuals, listw = spdep::nb2listw(spdep::poly2nb(nat_spatial)))
  cat("Moran's I for residuals (", nationality, "):", moran$estimate, "\n")
}
```

## Filtered years
```{r}
# Define the years to study
years_to_study <- c(2001, 2011, 2021) # Modify these years as needed

# Filter spatial data for selected years
filtered_data <- final_data_sf_sf %>%
  filter(Year %in% years_to_study) %>%
  left_join(london_data, by = "msoa11cd") %>%
  st_as_sf()

# Loop over nationalities
for (nationality in nationalities) {
  # Filter data for the current nationality and years
  nat_data <- filtered_data %>% filter(Nationality == nationality)
  
  # Ensure spatial features are valid
  nat_data <- nat_data %>% st_make_valid()
  
  # Convert to Spatial* object (required by spgwr)
  nat_spatial <- as(nat_data, "Spatial")
  
  # Bandwidth selection (adaptive GWR)
  bandwidth <- gwr.sel(as.formula(best_formula), data = nat_spatial, coords = coordinates(nat_spatial), gweight = gwr.Gauss)
  
  # Run GWR
  gwr_model <- gwr(as.formula(best_formula), data = nat_spatial, coords = coordinates(nat_spatial), bandwidth = bandwidth, gweight = gwr.Gauss, fit.points = TRUE)
  
  # Extract GWR results
  gwr_results <- as.data.frame(gwr_model$SDF)
  
  # Save results back to spatial data
  nat_data <- cbind(nat_data, gwr_results)
  
  # Save filtered GWR output to a shapefile
  st_write(nat_data, paste0("gwr_outputs/GWR_", nationality, "_FilteredYears.shp"), delete_dsn = TRUE)
  
  # Plot coefficient maps for selected predictors
  for (variable in c("HousingIndex", "RentPrice", "Tenancy", "LaggedSettlement")) {
    ggplot(nat_data) +
      geom_sf(aes_string(fill = variable), color = NA) +
      scale_fill_viridis_c() +
      labs(title = paste("GWR Coefficient Map for", variable, "(", nationality, " - Filtered Years)"), fill = "Coefficient") +
      theme_minimal() +
      ggsave(paste0("gwr_outputs/GWR_Coefficient_Filtered_", nationality, "_", variable, ".png"), width = 8, height = 6)
  }
}
```

# Diagnostico
```{r}
# Calculate RMSE for GWR
observed <- gwr_model$SDF$Dissimilarity_Index # Replace with your dependent variable's name
predicted <- gwr_model$SDF$gwr.pred           # GWR-predicted values
rmse <- sqrt(mean((observed - predicted)^2, na.rm = TRUE))

# Print and save RMSE
cat("RMSE for GWR (", nationality, "):", rmse, "\n")
write.table(data.frame(Nationality = nationality, RMSE = rmse),
            file = paste0("gwr_outputs/RMSE_", nationality, ".txt"), row.names = FALSE, append = TRUE)


# Additional diagnostics for each GWR model
for (nationality in nationalities) {
  # Filter data for the current nationality
  nat_data <- spatial_data %>% filter(Nationality == nationality)
  
  # Ensure spatial features are valid
  nat_data <- nat_data %>% st_make_valid()
  
  # Convert to Spatial* object (required by spgwr)
  nat_spatial <- as(nat_data, "Spatial")
  
  # Bandwidth selection (adaptive GWR)
  bandwidth <- gwr.sel(as.formula(best_formula), data = nat_spatial, coords = coordinates(nat_spatial), gweight = gwr.Gauss)
  
  # Run GWR
  gwr_model <- gwr(as.formula(best_formula), data = nat_spatial, coords = coordinates(nat_spatial), bandwidth = bandwidth, gweight = gwr.Gauss, fit.points = TRUE)
  
  # Extract GWR results
  gwr_results <- as.data.frame(gwr_model$SDF)
  local_r2 <- gwr_results$localR2
  residuals <- gwr_results$residuals
  
  # Save results back to spatial data
  nat_data <- cbind(nat_data, gwr_results)
  
  # Save GWR output to a shapefile
  st_write(nat_data, paste0("gwr_outputs/GWR_", nationality, ".shp"), delete_dsn = TRUE)
  
  # Plot coefficient maps for selected predictors
  for (variable in c("HousingIndex", "RentPrice", "Tenancy", "LaggedSettlement")) {
    ggplot(nat_data) +
      geom_sf(aes_string(fill = variable), color = NA) +
      scale_fill_viridis_c() +
      labs(title = paste("GWR Coefficient Map for", variable, "(", nationality, ")"), fill = "Coefficient") +
      theme_minimal() +
      ggsave(paste0("gwr_outputs/GWR_Coefficient_", nationality, "_", variable, ".png"), width = 8, height = 6)
  }
  
  # Plot local R-squared
  ggplot(nat_data) +
    geom_sf(aes(fill = local_r2), color = NA) +
    scale_fill_viridis_c() +
    labs(title = paste("Local R-squared Map (", nationality, ")"), fill = "R-squared") +
    theme_minimal() +
    ggsave(paste0("gwr_outputs/GWR_LocalR2_", nationality, ".png"), width = 8, height = 6)
  
  # Residual diagnostics
  ggplot(nat_data) +
    geom_sf(aes(fill = residuals), color = NA) +
    scale_fill_viridis_c() +
    labs(title = paste("Residual Map (", nationality, ")"), fill = "Residuals") +
    theme_minimal() +
    ggsave(paste0("gwr_outputs/GWR_Residuals_", nationality, ".png"), width = 8, height = 6)
  
  # Spatial autocorrelation of residuals (Moran's I)
  moran <- spdep::moran.test(residuals, listw = spdep::nb2listw(spdep::poly2nb(nat_spatial)))
  cat("Moran's I for residuals (", nationality, "):", moran$estimate, "\n")
}
```

##Temporal analysis of metrics
```{r}
# Define years to analyze
years <- unique(final_data_sf_sf$Year) # Extract unique years in the dataset

# Create a data frame to store temporal metrics
temporal_metrics <- data.frame()

# Loop over years and nationalities
for (year in years) {
  for (nationality in nationalities) {
    # Filter data for the current year and nationality
    nat_year_data <- spatial_data %>%
      filter(Nationality == nationality, Year == year)
    
    # Ensure spatial features are valid
    nat_year_data <- nat_year_data %>% st_make_valid()
    
    # Convert to Spatial* object
    nat_spatial <- as(nat_year_data, "Spatial")
    
    # Bandwidth selection
    bandwidth <- gwr.sel(as.formula(best_formula), data = nat_spatial, coords = coordinates(nat_spatial), gweight = gwr.Gauss)
    
    # Run GWR
    gwr_model <- gwr(as.formula(best_formula), data = nat_spatial, coords = coordinates(nat_spatial), bandwidth = bandwidth, gweight = gwr.Gauss, fit.points = TRUE)
    
    # Extract metrics
    observed <- gwr_model$SDF$Dissimilarity_Index # Replace with dependent variable
    predicted <- gwr_model$SDF$gwr.pred
    rmse <- sqrt(mean((observed - predicted)^2, na.rm = TRUE))
    mean_r2 <- mean(gwr_model$SDF$localR2, na.rm = TRUE) # Average local R-squared
    
    # Append metrics to the temporal data frame
    temporal_metrics <- rbind(temporal_metrics, data.frame(
      Year = year,
      Nationality = nationality,
      RMSE = rmse,
      Mean_R2 = mean_r2
    ))
  }
}

# Save temporal metrics to a CSV file
write.csv(temporal_metrics, "gwr_outputs/Temporal_Metrics.csv", row.names = FALSE)

# Visualize temporal trends
library(ggplot2)

# RMSE Trends
ggplot(temporal_metrics, aes(x = Year, y = RMSE, color = Nationality, group = Nationality)) +
  geom_line() +
  geom_point() +
  labs(title = "RMSE Trends Over Time", x = "Year", y = "RMSE") +
  theme_minimal() +
  ggsave("gwr_outputs/RMSE_Trends.png", width = 8, height = 6)

# Mean R-squared Trends
ggplot(temporal_metrics, aes(x = Year, y = Mean_R2, color = Nationality, group = Nationality)) +
  geom_line() +
  geom_point() +
  labs(title = "Mean R-squared Trends Over Time", x = "Year", y = "Mean R-squared") +
  theme_minimal() +
  ggsave("gwr_outputs/MeanR2_Trends.png", width = 8, height = 6)
```


