This document shows the code for the cleaning and joining of the demographic "census" data, and creation of its dataset. 

## Load necessary libraries
```{r}
install.packages(c("readxl", "dplyr", "tidyr", "stats", "readr", "zoo", 
                   "ggplot2", "sf", "tmap", "sp", "spdep", "MASS", 
                    "tidyverse", "stringr", "ineq", "FactoMineR", 
                    "factoextra", "reshape2"))

# Load libraries
library(readxl)        
library(dplyr)        
library(tidyr)         
library(stats)         
library(readr)         
library(zoo)           
library(ggplot2) 
library(sp)            
library(MASS)         
library(tidyverse)    
library(stringr)       
library(ineq)          
library(factoextra)  
library(reshape2)      
```

## Load census data
```{r}
setwd("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data")

# Demographic data
census_2001 <- read_csv("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/census 2001.csv")
census_2011 <- read_csv("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/census 2011 vf.csv")
census_2021 <- read_csv("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/census 2021.csv")

install.packages("sf", type = "binary")
library(sf)

# Define file paths for the shapefiles
msoa_path <- "/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/MSOA_2011_London_gen_MHW.shp"
lsoa_path <- "/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/LSOA_2011_London_gen_MHW.shp"
borough_path <- "/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/London_Borough_Excluding_MHW.shp"

# Read the shapefiles using st_read
msoa <- st_read(msoa_path, quiet = FALSE)
lsoa <- st_read(lsoa_path, quiet = FALSE)
borough <- st_read(borough_path, quiet = FALSE)
```

## Clean and join all data

### Spatial data
```{r}
# Eliminate columns
msoa <- dplyr::select(msoa, -USUALRES, -HHOLDRES, -COMESTRES, -POPDEN, -HHOLDS, -AVHHOLDSZ, -LAD11NM, -RGN11NM, -LAD11CD, -RGN11CD)

# Eliminate columns
lsoa <- dplyr::select(lsoa, -MSOA11CD, -MSOA11NM, -LAD11CD, -LAD11NM, -RGN11CD, -RGN11NM, -USUALRES, -HHOLDRES, -COMESTRES, -POPDEN, -HHOLDS, -AVHHOLDSZ)

# Eliminate columns
borough <- dplyr::select(borough, -HECTARES, -NONLD_AREA, -ONS_INNER, -SUB_2009, -SUB_2006)
```

### Demographic data

##### Census 2001
```{r}
# Load necessary packages
library(dplyr)

# Clean columns
names(census_2001)[names(census_2001) == "geography"] <- "MSOA11NM"
names(census_2001)[names(census_2001) == "geography code"] <- "MSOA11CD"
clean_2001 <- na.omit(census_2001)  
clean_2001$year <- 2001  

# Exclude unwanted columns
column_indices <- c(1, 2, 3, 4, 6, 7, 8, 9, 10, 11, 12, 13, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 32, 33, 34, 35, 36, 37, 38, 39, 40, 43, 44, 45, 46, 48, 49, 50, 51, 52, 55, 56, 57, 58, 60, 61, 62, 63, 64, 65, 66, 68, 69, 70, 71, 72, 73, 75, 76, 77, 78, 79, 80, 82, 83, 84, 85)
cleaner_2001 <- clean_2001[, column_indices]

# Clean the nationality names and convert to numeric codes
cleaner_2001 <- cleaner_2001 %>%
  rename_with(~ gsub("Country of Birth: ", "", .x))

# Identify nationality columns 
non_nationality_columns <- c("MSOA11CD", "MSOA11NM", "date", "All categories: Country of birth; measures: Value")
nationality_columns <- setdiff(names(cleaner_2001), non_nationality_columns)

# Ensure all nationality columns are numeric
cleaner_2001[nationality_columns] <- lapply(cleaner_2001[nationality_columns], as.numeric)

# Add majority nationality value column
cleaner_2001 <- cleaner_2001 %>%
  rowwise() %>%
  mutate(
    Majority_Nationality_Value = max(c_across(all_of(nationality_columns)))
  ) %>%
  ungroup()

# Select only the total population and specified nationalities
specific_nationalities <- c("Romania; measures: Value", "Poland; measures: Value", 
                            "Jamaica; measures: Value", "Other Caribbean and West Indies; measures: Value", 
                            "Somalia; measures: Value", "Kenya; measures: Value", 
                            "South Africa; measures: Value", "Zimbabwe; measures: Value", 
                            "Other  South and Eastern Africa; measures: Value", "Bangladesh; measures: Value", 
                            "South America; measures: Value")

# Column for total population
total_population_column <- "All categories: Country of birth; measures: Value"

# Eliminate columns
final_2001 <- cleaner_2001 %>%
  dplyr::select(MSOA11NM, MSOA11CD, date, Majority_Nationality_Value, total_population_column, all_of(specific_nationalities))

# Clean the column names
final_2001 <- final_2001 %>%
  rename_with(~ gsub("; measures: Value", "", .x))

# Create new columns for aggregated nationalities
library(dplyr)

final_2001 <- final_2001 %>%
  mutate(
    Caribbean = `Jamaica` + `Other Caribbean and West Indies`,
    SEAfrica = `Somalia` + `Kenya` + `South Africa` + `Zimbabwe` + `Other  South and Eastern Africa`,
    SouthAmerica = `South America`
  ) %>%
  dplyr::select(-`Jamaica`, -`Other Caribbean and West Indies`, -`Somalia`, -`Kenya`, -`South Africa`, -`Zimbabwe`, 
                -`Other  South and Eastern Africa`, -`South America`)

final_2001 <- final_2001 %>%
  rename(
    Total_Population_2001 = `All categories: Country of birth`,
    Majority_2001 = `Majority_Nationality_Value`,
    Caribbean_2001 = Caribbean,
    SEAfrica_2001 = SEAfrica,
    Poland_2001 = Poland,
    Romania_2001 = Romania,
    Bangladesh_2001 = Bangladesh,
    SouthAmerica_2001 = SouthAmerica
  )
```

##### Census 2011
```{r}
# Cleaning
names(census_2011)[names(census_2011) == "geography"] <- "MSOA11NM"
names(census_2011)[names(census_2011) == "geography code"] <- "MSOA11CD"
clean_2011 <- na.omit(census_2011)  
clean_2011$year <- 2011  

# Eliminate columns
column_indices <- c(1, 2, 3, 5, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 22, 23, 24, 25, 26, 27, 29, 30, 31, 32, 34, 35, 37, 39, 40, 41, 43, 44, 45, 46, 47, 48, 51, 52, 54, 55, 56, 58, 59, 60, 61, 62, 64, 65, 69, 70, 71, 72, 74, 75, 77, 79, 80, 81, 82)
cleaner_2011 <- clean_2011[, column_indices]

# Rename columns
specific_columns_corrected <- c(
  "Country of Birth: All categories: Country of birth; measures: Value",
  "Country of Birth: Europe: United Kingdom: England; measures: Value",
  "Country of Birth: Europe: Other Europe: EU countries: Accession countries April 2001 to March 2011: Romania; measures: Value",
  "Country of Birth: Europe: Other Europe: EU countries: Accession countries April 2001 to March 2011: Poland; measures: Value",
"Country of Birth: Africa: South and Eastern Africa: Kenya; measures: Value",                                                  "Country of Birth: Africa: South and Eastern Africa: Somalia; measures: Value",
"Country of Birth: Africa: South and Eastern Africa: South Africa; measures: Value",                                           "Country of Birth: Africa: South and Eastern Africa: Zimbabwe; measures: Value",                                               "Country of Birth: Africa: South and Eastern Africa: Other South and Eastern Africa; measures: Value", 
  "Country of Birth: Middle East and Asia: Southern Asia: Bangladesh; measures: Value",
  "Country of Birth: The Americas and the Caribbean: South America; measures: Value",
"Country of Birth: The Americas and the Caribbean: The Caribbean: Jamaica; measures: Value",                                                        "Country of Birth: The Americas and the Caribbean: The Caribbean: Other Caribbean; measures: Value" 
)

specific_nationalities <- c("total population", "Majority", "Romania", "Poland", "Kenya", "Somalia", "South Africa", "Zimbabwe", "Other South and Eastern Africa", "Bangladesh", "South America", "Jamaica", "Other Caribbean")

# Create final dataset
final_2011 <- cleaner_2011 %>%
  dplyr::select(MSOA11CD, MSOA11NM, date, all_of(specific_columns_corrected)) %>%
  dplyr::rename_with(~ specific_nationalities, .cols = all_of(specific_columns_corrected))

final_2011 <- final_2011 %>%
  dplyr::mutate(
    Caribbean = `Jamaica` + `Other Caribbean`,
    SEAfrica = `Somalia` + `Kenya` + `South Africa` + `Zimbabwe` + `Other South and Eastern Africa`
  ) %>%
  dplyr::select(-`Jamaica`, -`Other Caribbean`, -`Somalia`, -`Kenya`, -`South Africa`, -`Zimbabwe`, -`Other South and Eastern Africa`)

final_2011 <- final_2011 %>%
  rename(
    Total_Population_2011 = `total population`,
    Majority_2011 = `Majority`,
    Caribbean_2011 = Caribbean,
    SEAfrica_2011 = SEAfrica,
    Poland_2011 = Poland,
    Romania_2011 = Romania,
    Bangladesh_2011 = Bangladesh,
    SouthAmerica_2011 = `South America`
  )
```

##### Census 2021

```{r}
# Initial cleaning
names(census_2021)[names(census_2021) == "Middle layer Super Output Areas"] <- "MSOA11NM"
names(census_2021)[names(census_2021) == "Middle layer Super Output Areas Code"] <- "MSOA11CD"
clean_2021 <- na.omit(census_2021)

# Pivot wider
pivoted_2021 <- clean_2021 %>%
  pivot_wider(
    id_cols = c(MSOA11CD, MSOA11NM),
    names_from = `Country of birth (60 categories)`,
    values_from = Observation,
    values_fill = list(Observation = 0)
  )

clean_pivoted_2021 <- na.omit(pivoted_2021)
clean_pivoted_2021$date <- 2021
head(clean_pivoted_2021)

# Calculate total population from all nationality columns
total_columns <- grep("Europe|Africa|Middle East and Asia|The Americas and the Caribbean", names(clean_pivoted_2021), value = TRUE)
clean_pivoted_2021$Total_Population <- rowSums(clean_pivoted_2021[, total_columns], na.rm = TRUE)

# Define specific nationalities and their columns 
specific_columns <- c(
  "Europe: United Kingdom: England",
  "Europe: United Kingdom: Northern Ireland",
  "Europe: United Kingdom: Scotland",
  "Europe: United Kingdom: Wales",
  "Europe: Other Europe: EU countries: Countries that joined the EU between April 2001 and March 2011: Romania",
  "Europe: Other Europe: EU countries: Countries that joined the EU between April 2001 and March 2011: Poland",
  "Africa: South and Eastern Africa: Somalia",
  "Africa: South and Eastern Africa: Kenya",
  "Africa: South and Eastern Africa: South Africa",
  "Africa: South and Eastern Africa: Zimbabwe",
  "Africa: South and Eastern Africa: Other South and Eastern Africa",
  "Middle East and Asia: Southern Asia: Bangladesh",
  "The Americas and the Caribbean: South America: All South American countries",
  "The Americas and the Caribbean: The Caribbean: Jamaica",
  "The Americas and the Caribbean: The Caribbean: Other Caribbean"
)

# Prepare the final dataset
final_2021 <- clean_pivoted_2021 %>%
  dplyr::select(MSOA11CD, MSOA11NM, date, Total_Population, all_of(specific_columns)) %>%
  dplyr::rename(
    England_2021 = `Europe: United Kingdom: England`,
    Irland_2021 = `Europe: United Kingdom: Northern Ireland`,
    Scotland_2021 = `Europe: United Kingdom: Scotland`,
    Wales_2021 = `Europe: United Kingdom: Wales`,
    Romania_2021 = `Europe: Other Europe: EU countries: Countries that joined the EU between April 2001 and March 2011: Romania`,
    Poland_2021 = `Europe: Other Europe: EU countries: Countries that joined the EU between April 2001 and March 2011: Poland`,
    Somalia_2021 = `Africa: South and Eastern Africa: Somalia`,
    SEAfrica_2021 = `Africa: South and Eastern Africa: Other South and Eastern Africa`,
    Bangladesh_2021 = `Middle East and Asia: Southern Asia: Bangladesh`,
    SouthAmerica_2021 = `The Americas and the Caribbean: South America: All South American countries`,
    Kenya_2021 = `Africa: South and Eastern Africa: Kenya`,
    Zimbabwe_2021 = `Africa: South and Eastern Africa: Zimbabwe`,
    SAfrica_2021 = `Africa: South and Eastern Africa: South Africa`,
    Jamaica_2021 = `The Americas and the Caribbean: The Caribbean: Jamaica`,
    Caribbean_2021 = `The Americas and the Caribbean: The Caribbean: Other Caribbean`
  )

final_2021 <- final_2021 %>%
  dplyr::mutate(
    Caribbean_2021 = `Jamaica_2021` + `Caribbean_2021`,
    SEAfrica_2021 = `Somalia_2021` + `Kenya_2021` + `SAfrica_2021` + `Zimbabwe_2021` + `SEAfrica_2021`,
    Majority_2021 = `England_2021` + `Irland_2021` + `Scotland_2021` + `Wales_2021`
  ) %>%
  dplyr::select(-`Jamaica_2021`, -`Somalia_2021`, -`Kenya_2021`, -`SAfrica_2021`, -`Zimbabwe_2021`)

final_2021 <- final_2021 %>%
  dplyr::select(-England_2021, -Irland_2021, -Scotland_2021, -Wales_2021)
```

#### Create a complete census dataset
```{r}
# Merge the datasets by MSOA codes and names
combined_census <- final_2001 %>%
  full_join(final_2011, by = c("MSOA11CD", "MSOA11NM")) %>%
  full_join(final_2021, by = c("MSOA11CD", "MSOA11NM"))

# Clean NA values
final_census <- na.omit(combined_census)

# Eliminate columns
final_census <- dplyr::select(final_census, -date, -date.x, -date.y)

final_census <- final_census %>%
  rename(Total_Population_2021 = Total_Population)
```

#### Interpolate demogrpahic data set for non-census years
```{r}

colnames(final_census)
names(final_census)[17] <- "Total_Population_2021"
colnames(final_census)
names(final_census)[17] <- "Caribbean_2011"
names(final_census)[19] <- "Total_Population_2021"

# Transform the data into a long format with separate Nationality and Year columns
data_long <- final_census %>%
  pivot_longer(
    cols = -c(MSOA11CD, MSOA11NM), 
    names_to = c("Nationality", "Year"), 
    names_pattern = "(.*)_(\\d{4})",
    values_to = "Population"
  )

# Convert Year to numeric
data_long$Year <- as.numeric(data_long$Year)

# Interpolate for odd years between 2001 and 2021
library(zoo)
data_long_interpolated <- data_long %>%
  group_by(MSOA11CD, Nationality) %>%
  complete(Year = seq(2001, 2021, by = 1)) %>%
  arrange(Year) %>%
  mutate(Population = na.approx(Population, rule = 2))

# Filter to keep only odd years
data_long_odd_years <- data_long_interpolated %>%
  filter(Year %% 2 == 1)

```
#### Join geometry to final dataset

```{r}
library(tidyr)

# Fill down the MSOA11NM column
data_long_odd_years <- data_long_odd_years %>%
  group_by(MSOA11CD) %>%
  fill(MSOA11NM, .direction = "downup") %>%
  ungroup()

# Join to msoa shp
msoa_census <- left_join(msoa, data_long_odd_years, by = c("MSOA11NM", "MSOA11CD"))
```

#### Interpolate demogrpahic data set for non-census years
```{r}
# Calculate proportions of each nationality
census_final <- msoa_census %>%
  group_by(MSOA11CD, Year) %>%
  mutate(
    Total_Pop = Population[Nationality == "Total_Population"],
    Proportion = ifelse(Nationality == "Total_Population", 1, Population / Total_Pop)
  ) %>%
  ungroup()
```

## Calculation of 3 Segregation analysis

```{r}
# Replace zero proportions with a small value
census_final$Proportion <- ifelse(census_final$Proportion == 0, 0.00001, census_final$Proportion)
```

### Dissimilarity index

#### Dissimiliarity index

```{r}
library(tidyverse)
# Reshape dataset to wide format
wide_census <- census_final %>%
  pivot_wider(
    names_from = Nationality, 
    values_from = c(Population, Proportion)
  )

library(dplyr)

# Drop specific columns
wide_census <- wide_census %>%
  select(-Population_Total_Population, -Population_NA, -Proportion_Total_Population, -Proportion_NA)

library(dplyr)

# Function to calculate Dissimilarity Index for one nationality
calculate_diss_index <- function(minority_proportion, majority_proportion) {
  0.5 * abs(minority_proportion - majority_proportion)
}

# Calculate the Dissimilarity Index for each nationality
wide_census<- wide_census %>%
  rowwise() %>%
  mutate(
    Diss_Index_Bangladesh = calculate_diss_index(Proportion_Bangladesh, Proportion_Majority),
    Diss_Index_Caribbean = calculate_diss_index(Proportion_Caribbean, Proportion_Majority),
    Diss_Index_Poland = calculate_diss_index(Proportion_Poland, Proportion_Majority),
    Diss_Index_Romania = calculate_diss_index(Proportion_Romania, Proportion_Majority),
    Diss_Index_SEAfrica = calculate_diss_index(Proportion_SEAfrica, Proportion_Majority),
    Diss_Index_SouthAmerica = calculate_diss_index(Proportion_SouthAmerica, Proportion_Majority)
  ) %>%
  ungroup() 

# Drop specific columns
wide_census <- wide_census %>%
  select(-Dissimilarity_Index)
```

#### Lagged diss (-1)

```{r}
# Calculate the lagged dissimilarity index for each nationality
wide_census<- wide_census %>%
  arrange(MSOA11CD, Year) %>%  # Ensure data is sorted by MSOA and Year
  group_by(MSOA11CD) %>%       # Group by MSOA (no Nationality needed in wide format)
  mutate(
    lagged_diss_Bangladesh = if_else(Year == 2001, Diss_Index_Bangladesh, lag(Diss_Index_Bangladesh)),
    lagged_diss_Caribbean = if_else(Year == 2001, Diss_Index_Caribbean, lag(Diss_Index_Caribbean)),
    lagged_diss_Poland = if_else(Year == 2001, Diss_Index_Poland, lag(Diss_Index_Poland)),
    lagged_diss_Romania = if_else(Year == 2001, Diss_Index_Romania, lag(Diss_Index_Romania)),
    lagged_diss_SEAfrica = if_else(Year == 2001, Diss_Index_SEAfrica, lag(Diss_Index_SEAfrica)),
    lagged_diss_SouthAmerica = if_else(Year == 2001, Diss_Index_SouthAmerica, lag(Diss_Index_SouthAmerica))
  ) %>%
  ungroup()  

# View the updated dataset with lagged columns
print(wide_census %>% select(MSOA11CD, Year, starts_with("lagged_diss")))
```

### Isolation and Thiels H index

```{r}
library(dplyr)
library(tidyr)

# Calculate Theil's H and Isolation Index
wide_census <- wide_census %>%
  rowwise() %>%
  mutate(
    # Total number of groups (including Majority and all nationalities)
    k = 7,  # Adjust this if you have more or fewer groups
    
    # Theil's H calculation for each nationality
    Theil_H_Bangladesh = -Proportion_Bangladesh * log(Proportion_Bangladesh) / log(k),
    Theil_H_Caribbean = -Proportion_Caribbean * log(Proportion_Caribbean) / log(k),
    Theil_H_Poland = -Proportion_Poland * log(Proportion_Poland) / log(k),
    Theil_H_Romania = -Proportion_Romania * log(Proportion_Romania) / log(k),
    Theil_H_SEAfrica = -Proportion_SEAfrica * log(Proportion_SEAfrica) / log(k),
    Theil_H_SouthAmerica = -Proportion_SouthAmerica * log(Proportion_SouthAmerica) / log(k),
    
    # Isolation Index calculation for each nationality
    Isolation_Bangladesh = Proportion_Bangladesh * (Population_Bangladesh / Total_Pop),
    Isolation_Caribbean = Proportion_Caribbean * (Population_Caribbean / Total_Pop),
    Isolation_Poland = Proportion_Poland * (Population_Poland / Total_Pop),
    Isolation_Romania = Proportion_Romania * (Population_Romania / Total_Pop),
    Isolation_SEAfrica = Proportion_SEAfrica * (Population_SEAfrica / Total_Pop),
    Isolation_SouthAmerica = Proportion_SouthAmerica * (Population_SouthAmerica / Total_Pop)
  ) %>%
  ungroup()
```

```{r}
wide_census_cleaned <- wide_census %>%
  select(
    -starts_with("Population_"),  
    -starts_with("Majority")     
  )

wide_census_cleaner <- wide_census_cleaned %>%
  select(
    -starts_with("Total_Pop")
  )

wide_census_cleanerer <- wide_census_cleaner %>%
  select(
    -starts_with("k")
  )

final_long_census <- wide_census_cleanerer %>%
  pivot_longer(
    cols = starts_with("Proportion") | starts_with("Diss_Index") | starts_with("lagged_diss") | starts_with("Theil_H") | starts_with("Isolation"),
    names_to = c("Metric", "Nationality"),
    names_pattern = "(.*)_(.*)",  # Split at the last underscore
    values_to = "Value"
  ) %>%
  pivot_wider(
    names_from = "Metric",
    values_from = "Value"
  )

# Remove rows where Nationality is "Majority"
final_long_census <- final_long_census %>%
  filter(Nationality != "Majority")
```

## Download CSV
```{r}
# Remove the geometry column from the data frame
census_no_geo <- st_drop_geometry(final_long_census)

# Download as a csv 
write.csv(census_no_geo, "census_data.csv", row.names = FALSE)
```




