This document shows the process of cleaning and joining the datasets of housing and neighborhood data.  

## Load necessary libraries and data
```{r}
install.packages("readr")  # Install the readr package
library(readr) 
install.packages("dplyr")
library(dplyr)
install.packages("sf")
library(sf)

setwd("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data")

# Load datasets
house_prices <- read_csv("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/House prices.csv")
accomodation_2011 <- read_csv("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/housing 2011.csv")
accomodation_2021 <- read_csv("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/housing 2021.csv")
tenure_2001 <- read_csv("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/home tenure 2001.csv")
tenure_2011 <- read_csv("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/home tenure 2011.csv")
tenure_2021 <- read_csv("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/home tenure 2021.csv")
private_rents <- read_csv("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/new rent.csv")
affordable_budget <- read_csv("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/affordable budget.csv")
price_london <- read_csv("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/price_london.csv")
tenure_FINAL <- read_csv("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/tenure_FINAL.csv")


# Spatial data
msoa <- st_read("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/MSOA_2011_London_gen_MHW.shp")
lsoa <- st_read("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/LSOA_2011_London_gen_MHW.shp")
borough <- st_read("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/London_Borough_Excluding_MHW.shp")
```

## Cleaning and joining of Housing data

### Houseprice data
```{r}
# Rename columns to keep only the last year in the range
colnames(house_prices) <- gsub("Year ending (Dec|Mar|Jun|Sep) ", "", colnames(house_prices))
colnames(house_prices)[colnames(house_prices) == "Local authority code"] <- "LAD11CD"
colnames(house_prices)[colnames(house_prices) == "Local authority name"] <- "LAD11NM"

# Eliminate all columns before 2001
years_to_keep <- as.character(seq(2001, 2021, by = 2))
columns_to_keep <- c("LAD11CD", "LAD11NM", years_to_keep)
house_prices <- house_prices %>%
  dplyr::select(all_of(columns_to_keep))

# Make all values integers
house_prices[ , 3:ncol(house_prices)] <- lapply(house_prices[ , 3:ncol(house_prices)], function(x) as.integer(gsub(",", "", x)))

# Calculate at MSOA level
msoa_houseprice <- msoa %>%
  left_join(house_prices, by = "LAD11NM")

# Select relevant columns
msoa_houseprice <- msoa_houseprice %>%
  dplyr::select(MSOA11CD, MSOA11NM, all_of(years_to_keep), geometry)

# Convert the geometry column to an sf object
msoa_houseprice <- st_as_sf(msoa_houseprice, crs = 27700)

#Calculate average msoa houseprice
average_houseprice <- msoa_houseprice %>%
  dplyr::select(-geometry) %>%  # Exclude the geometry column
  group_by(MSOA11CD, MSOA11NM) %>%
  summarise(across(starts_with("20"), ~ mean(.x, na.rm = TRUE), .names = "avg_{.col}")) %>%
  ungroup()

install.packages("tidyr")
library(tidyr)

long_format_houseprice <- average_houseprice %>%
  pivot_longer(
    cols = starts_with("avg_"),   
    names_to = "Year",            
    values_to = "Houseprice"      
  )

# Remove the "avg_" prefix from the Year column
long_format_houseprice <- long_format_houseprice %>%
  mutate(Year = gsub("avg_", "", Year))

# Convert the Year column to numeric if needed
long_format_houseprice <- long_format_houseprice %>%
  mutate(Year = as.numeric(Year))

# Incorporate housing london as a new row
library(lubridate)

# Parse the Period column and prepare London_price data
processed_london_price <- price_london %>%
  mutate(
    Period = as.Date(paste0(Period, "-01")),                
    Year = year(Period),                                   
    `Average price All property types` = as.numeric(`Average price All property types`)  
  ) %>%
  group_by(Year) %>%
  summarise(
    Houseprice = mean(`Average price All property types`, na.rm = TRUE)  
  ) %>%
  complete(Year = 2001:2021) %>%                            
  fill(Houseprice, .direction = "downup") %>%              
  filter(!is.na(Houseprice)) %>%                           
  mutate(
    MSOA11CD = "London",                                   
    MSOA11NM = "London",
    geometry = NA                                         
  ) %>%
  dplyr::select(MSOA11CD, MSOA11NM, Year, Houseprice, geometry)   

processed_london_price <- processed_london_price %>% filter(Year >= 2001 & Year <= 2021)
processed_london_price <- processed_london_price %>%
  mutate(geometry = st_sfc(st_geometrycollection(), crs = st_crs(long_format_houseprice))) %>%
  st_as_sf(crs = st_crs(long_format_houseprice))

# Combine with long_format_houseprice
combined_houseprice <- long_format_houseprice %>%
  bind_rows(processed_london_price)
```

### Occupation
```{r}
# Accommodation 2011

# Select columns 'Area' and 'Overcrowded ( -1 or less)' from 'house_2011'
accomodation_2011 <- accomodation_2011[, c('Area', 'Overcrowded ( -1 or less)')]

# Calculate total population per borough
msoa <- msoa %>%
  group_by(LAD11NM) %>%
  mutate(TotalPopulation = sum(USUALRES))

# Calculate the proportion for each MSOA
msoa <- msoa %>%
  mutate(Proportion = USUALRES / TotalPopulation)

# Merge accommodation2011 data with MSOA data
msoa_accomodation_2011 <- merge(msoa, accomodation_2011, by.x = "LAD11NM", by.y = "Area")

# Calculate distributed data
msoa_accomodation_2011 <- msoa_accomodation_2011 %>%
  mutate(DistributedOvercrowded = Proportion * `Overcrowded ( -1 or less)`)

# Select final columns
final_accomodation_2011 <- msoa_accomodation_2011 %>%
  dplyr::select(MSOA11NM, MSOA11CD, geometry, DistributedOvercrowded)

final_accomodation_2011 <- final_accomodation_2011 %>%
  rename(
    occupation_2011 = DistributedOvercrowded
  )

# Accommodation 2021

# Selecting columns for accommodation_2021 dataframe
accomodation_2021 <- accomodation_2021[, c('Lower layer Super Output Areas Code',
                             'Lower layer Super Output Areas',
                             'Occupancy rating for bedrooms (5 categories)', 'Observation')]

# Check the structure of the dataframe to ensure column names are correct
str(accomodation_2021)

# Filter the rows with occupancy rating of bedrooms: "-1 or less"
clean_accomodation_2021 <- accomodation_2021 %>%
  group_by(`Lower layer Super Output Areas Code`) %>%
  filter(row_number() %in% 33:40) %>%
  ungroup()

# Summarize the total values for each "Lower layer Super Output Areas Code"
aggregated_accomodation_2021 <- clean_accomodation_2021 %>%
  group_by(`Lower layer Super Output Areas Code`, `Lower layer Super Output Areas`, `Occupancy rating for bedrooms (5 categories)`) %>%
  summarise(Total_Observations = sum(Observation, na.rm = TRUE)) %>%
  ungroup()

# Merge to the LSOA codes
merged_accomodation_2021 <- merge(aggregated_accomodation_2021, lsoa, by.x = "Lower layer Super Output Areas Code", by.y = "LSOA11CD")

# Aggregate at MSOA level
agg_accomodation_2021 <- merged_accomodation_2021 %>%
  group_by(MSOA11CD, MSOA11NM) %>%
  summarise(
    Total_Observations = sum(Total_Observations),
    USUALRES = sum(USUALRES),
    HHOLDRES = sum(HHOLDRES),
    COMESTRES = sum(COMESTRES),
    POPDEN = mean(POPDEN),
    HHOLDS = sum(HHOLDS),
    AVHHOLDSZ = mean(AVHHOLDSZ)
  )

# Select final columns
final_accomodation_2021 <- agg_accomodation_2021 %>%
  dplyr::select(MSOA11NM, MSOA11CD, 'Total_Observations')

# Merge to the MSOA codes
ffinal_accomodation_2021 <- merge(final_accomodation_2021, msoa, by = "MSOA11NM")

# Select final columns
fffinal_accomodation_2021 <- ffinal_accomodation_2021 %>%
  dplyr::select(MSOA11NM, MSOA11CD.x, 'Total_Observations', geometry)

fffinal_accomodation_2021 <- fffinal_accomodation_2021 %>%
  rename(
    occupation_2021 = 'Total_Observations'
  )

# Total accommodation dataset

# Join both datasets
accommodation <- final_accomodation_2011 %>%
  inner_join(fffinal_accomodation_2021, by = c("MSOA11NM", "MSOA11CD" = "MSOA11CD.x", "geometry"))

# Interpolate missing years
# Convert the geometry column to an sf object
accommodation <- st_as_sf(accommodation, crs = 27700)

# Reshape the data for easier interpolation
accommodation_long <- accommodation %>%
  pivot_longer(
    cols = c(occupation_2011, occupation_2021), 
    names_to = "Year", 
    names_prefix = "occupation_", 
    values_to = "Occupation"
  ) %>%
  mutate(Year = as.numeric(Year))

# Define years to interpolate
years <- seq(2001, 2021, by = 2)

# Perform interpolation for each MSOA
interpolated_accommodation <- accommodation_long %>%
  group_by(MSOA11NM, MSOA11CD, geometry) %>%
  do({
    data.frame(Year = years,
               Occupation = approx(.$Year, .$Occupation, xout = years, rule = 2)$y)
  }) %>%
  ungroup() %>%
  st_as_sf()

# Calculate the average occupation per year for London
london_accommodation <- interpolated_accommodation %>%
  st_drop_geometry() %>%              
  group_by(Year) %>%                   
  summarise(
    Occupation = mean(Occupation, na.rm = TRUE)  
  ) %>%
  mutate(
    MSOA11NM = "London",               
    MSOA11CD = "London"                
  )

# Add an empty geometry column for London
london_accommodation$geometry <- st_sfc(
  rep(st_geometrycollection(), nrow(london_accommodation)),  
  crs = st_crs(interpolated_accommodation)
)

# Convert London rows to sf
london_accommodation <- st_as_sf(london_accommodation, crs = st_crs(interpolated_accommodation))

# Append London rows to the original dataset
accommodation_with_london <- bind_rows(interpolated_accommodation, london_accommodation)
```

### Tenancy data
```{r}
# Calculate proportions 2001
tenure_2001 <- tenure_2001 %>%
  mutate(`2001 Owned` = `Tenure: Owned: Total; measures: Value` / `Tenure: All categories: Tenure; measures: Value`,
         `2001 Rented` = `Tenure: Private rented: Total; measures: Value` / `Tenure: All categories: Tenure; measures: Value`,
         `2001 Social` = `Tenure: Social rented: Total; measures: Value` / `Tenure: All categories: Tenure; measures: Value`)

# Keep only the required columns 2001
tenure_2001 <- tenure_2001 %>%
  dplyr::select(`geography code`, `2001 Owned`, `2001 Rented`, `2001 Social`)

# Ensure columns are numeric 2011
tenure_2011 <- tenure_2011 %>%
  mutate(
    `Owned outright` = as.numeric(`Owned outright`),
    `Owned with a mortgage or loan` = as.numeric(`Owned with a mortgage or loan`),
    `Shared ownership` = as.numeric(`Shared ownership`),
    `Rented from Local Authority` = as.numeric(`Rented from Local Authority`),
    `Other social rented` = as.numeric(`Other social rented`),
    `Private landlord or letting agency` = as.numeric(`Private landlord or letting agency`),
    `Other private rented` = as.numeric(`Other private rented`),
    `All Households` = as.numeric(`All Households`)
  )

# Perform the mutate operations
tenure_2011 <- tenure_2011 %>%
  mutate(
    `2011 Owned` = (`Owned outright` + `Owned with a mortgage or loan` + `Shared ownership`) / `All Households`,
    `2011 Social` = (`Rented from Local Authority` + `Other social rented`) / `All Households`,
    `2011 Rented` = (`Private landlord or letting agency` + `Other private rented`) / `All Households`
  )

# Keep only the required columns 2011
tenure_2011 <- tenure_2011 %>%
  dplyr::select(`LSOA code`, `2011 Owned`, `2011 Rented`, `2011 Social`)

# Ensure columns are numeric 2021
tenure_2021 <- tenure_2021 %>%
  mutate(
    `Owned outright` = as.numeric(`Owned outright`),
    `Owned with a mortgage or loan` = as.numeric(`Owned with a mortgage or loan`),
    `Shared ownership` = as.numeric(`Shared ownership`),
    `Rented from Local Authority` = as.numeric(`Rented from Local Authority`),
    `Other social rented` = as.numeric(`Other social rented`),
    `Private landlord or letting agency` = as.numeric(`Private landlord or letting agency`),
    `Other private rented` = as.numeric(`Other private rented`),
    `All Households` = as.numeric(`All Households`)
  )

# Perform the mutate operations
tenure_2021 <- tenure_2021 %>%
  mutate(
    `2021 Owned` = (`Owned outright` + `Owned with a mortgage or loan` + `Shared ownership`) / `All Households`,
    `2021 Social` = (`Rented from Local Authority` + `Other social rented`) / `All Households`,
    `2021 Rented` = (`Private landlord or letting agency` + `Other private rented`) / `All Households`
  )

# Keep only the required columns 2021
tenure_2021 <- tenure_2021 %>%
  dplyr::select(`LSOA code`, `2021 Owned`, `2021 Rented`, `2021 Social`)

# Join the datasets
tenure_2001 <- tenure_2001 %>%
  rename('LSOA code' = `geography code`)
tenure_2001 <- tenure_2001 %>%
  mutate(`LSOA code` = as.character(`LSOA code`))
tenure_2011 <- tenure_2011 %>%
  mutate(`LSOA code` = as.character(`LSOA code`))
tenure_2021 <- tenure_2021 %>%
  mutate(`LSOA code` = as.character(`LSOA code`))
joined_tenure <- tenure_2001 %>%
  full_join(tenure_2011, by = "LSOA code") %>%
  full_join(tenure_2021, by = "LSOA code")
cleaned_joined_tenure <- joined_tenure %>%
  drop_na()

# Interpolate missing years
years <- seq(2001, 2021, by = 2)
# Process the data
interpolated_tenure <- cleaned_joined_tenure %>%
  pivot_longer(cols = -`LSOA code`, names_to = "variable", values_to = "value") %>%
  separate(variable, into = c("year", "category"), sep = " ") %>%
  mutate(year = as.numeric(year)) %>%
  complete(`LSOA code`, category, year = years) %>%
  group_by(`LSOA code`, category) %>%
  arrange(year) %>%
  mutate(value = zoo::na.approx(value, maxgap = Inf, na.rm = FALSE)) 

# Calculate at msoa level
lsoa_tenure <- lsoa %>%
  left_join(interpolated_tenure, by = c("LSOA11CD" = "LSOA code"))
# Select the required columns
selected_tenure <- lsoa_tenure %>%
  dplyr::select(MSOA11CD, category, year, value)

wide_tenure <- selected_tenure %>%
  pivot_wider(names_from = category,
values_from = value)

# get unique msoa values
library(dplyr)
# Grouping by MSOA and year, then calculating average tenure (Owned, Rented, Social)
grouped_data <- wide_tenure %>%
  group_by(MSOA11CD, year) %>%
  summarise(
    average_owned = mean(Owned, na.rm = TRUE),
    average_rented = mean(Rented, na.rm = TRUE),
    average_social = mean(Social, na.rm = TRUE)
  )

# Add London data
library(stringr)

# Rename columns 1 and 2
tenure_FINAL <- tenure_FINAL %>%
  rename(
    MSOA11CD = 1,
    MSOA11NM = 2
  )

tenure_FINAL <- tenure_FINAL %>%
  mutate(across(
    -c(MSOA11CD, MSOA11NM), 
    ~ as.numeric(gsub("[^0-9.]", "", .)), 
    .names = "cleaned_{col}"
  ))

# Ensure all relevant columns have the same data type
tenure_FINAL <- tenure_FINAL %>%
  mutate(across(matches("^\\d{4}_"), as.numeric)) 

# Reshape the data to long format
tenure_LONG <- tenure_FINAL %>%
  pivot_longer(
    cols = matches("^\\d{4}_"), 
    names_to = "Category_Year", 
    values_to = "Value" 
  ) %>%
  separate(
    col = Category_Year, 
    into = c("Year", "Category"), 
    sep = "_", 
    convert = TRUE 
  )

# Calculate averages
tenure_LONG <- tenure_LONG %>%
  group_by(MSOA11CD, MSOA11NM, Year) %>%
  summarise(
    Total = sum(Value[Category == "Total"]),
    average_owned = sum(Value[Category == "Own Outright"] + Value[Category == "Buying with mortgage"]) / Total,
    average_rented = sum(Value[Category == "Rented from Private landlord"]) / Total,
    average_social = sum(Value[Category == "Rented from Local Authority or Housing Association"]) / Total
  ) %>%
  ungroup()

# Filter to keep only row 34 (London information)
tenure_FINAL_long_filtered <- tenure_LONG %>%
  filter(MSOA11CD == "E12000007")

# Keep only necessary columns
tenure_FINAL_long_filtered <- tenure_FINAL_long_filtered %>%
  dplyr::select(MSOA11CD, MSOA11NM, Year, average_owned, average_rented, average_social)

# Define the complete range of years
all_years <- 2001:2021

# Interpolate the missing years for each average column
interpolated_tenure <- tenure_FINAL_long_filtered %>%
  group_by(MSOA11CD, MSOA11NM) %>%                
  complete(Year = all_years) %>%                  
  arrange(Year) %>%                               
  mutate(
    average_owned = if (sum(!is.na(average_owned)) > 1) {
      approx(x = Year[!is.na(average_owned)], 
             y = average_owned[!is.na(average_owned)], 
             xout = Year, 
             rule = 2)$y
    } else {
      average_owned
    },
    average_rented = if (sum(!is.na(average_rented)) > 1) {
      approx(x = Year[!is.na(average_rented)], 
             y = average_rented[!is.na(average_rented)], 
             xout = Year, 
             rule = 2)$y
    } else {
      average_rented
    },
    average_social = if (sum(!is.na(average_social)) > 1) {
      approx(x = Year[!is.na(average_social)], 
             y = average_social[!is.na(average_social)], 
             xout = Year, 
             rule = 2)$y
    } else {
      average_social
    }
  ) %>%
  ungroup()
```

### Social Budget data
```{r}
# Filter rows with MSOA codes starting with "E090000..." or "E12000007"
affordable_budget <- affordable_budget %>%
  filter(grepl("^E090000", `Current\nONS code`) | `Current\nONS code` == "E12000007")

# Rename the "E12000007" row to "london_budget"
affordable_budget <- affordable_budget %>%
  mutate(`Current\nONS code` = if_else(`Current\nONS code` == "E12000007", "london_budget", `Current\nONS code`))

# Clean empty row
affordable_budget <- affordable_budget %>%
  filter(row_number() != 1)

# Eliminate unnecessary columns
affordable_budget <- affordable_budget %>%
  dplyr::select(-1, -3, -4, -5, -6, -7, -8, -9, -10, -11, -12, -34, -35)

# Rename the first column
colnames(affordable_budget)[1] <- "LAD11CD"

# Extract column names
col_names <- colnames(affordable_budget)

# Modify column names to keep only the last two digits for the time range columns
new_col_names <- sapply(col_names, function(x) {
  if (grepl("^[0-9]{4}-[0-9]{2}$", x)) {
    return(str_sub(x, -2))
  } else {
    return(x)
  }
})
colnames(affordable_budget) <- new_col_names

# Modify column names to start with "20" for the time range columns
col_names <- colnames(affordable_budget)
new_col_names <- sapply(col_names, function(x) {
  if (grepl("^[0-9]{2}$", x)) {
    return(paste0("20", x))
  } else {
    return(x)
  }
})
colnames(affordable_budget) <- new_col_names

# Convert to long format
budget_long <- affordable_budget %>%
  pivot_longer(
    cols = starts_with("20"), 
    names_to = "Year",        
    values_to = "Budget"      
  )

# Filter out rows where the year is even
budget_filtered <- budget_long %>%
  filter(as.integer(Year) %% 2 != 0)

budget_lad11cd <- budget_filtered %>%
  full_join(msoa, by = "LAD11CD")

budget_msoa <- budget_lad11cd %>%
  mutate(
    budget_proportion = if_else(
      MSOA11CD == "london_budget", 
      Budget,                      
      Budget * Proportion          
    )
  )

budget_msoa <- budget_msoa %>%
  mutate(
    budget_proportion = if_else(
      LAD11CD == "london_budget",   
      as.numeric(Budget),          
      as.numeric(Budget) * Proportion 
    )
  )

budget_msoa <- budget_msoa %>%
  mutate(
    MSOA11CD = if_else(LAD11CD == "london_budget", "London", MSOA11CD), # Rename MSOA11CD for London_budget
    MSOA11NM = if_else(LAD11CD == "london_budget", "London", MSOA11NM)  # Rename MSOA11NM for London_budget
  )

# Select the desired columns
final_budget <- budget_msoa %>%
  dplyr::select(MSOA11CD, MSOA11NM, geometry, Year, budget_proportion)
```

### Rent data
```{r}
rents_cleaned <- private_rents[-c(1, 2), ]

# Transpose rows 3 and 4 to merge them
headers <- paste(rents_cleaned[1, ], rents_cleaned[2, ], sep = " ")
headers <- gsub("^\\s+|\\s+$", "", headers) # Trim leading/trailing whitespace
headers <- gsub("NA", "", headers)         # Remove any "NA" from merged headers

# Assign the new headers to the dataset
colnames(rents_cleaned) <- headers
rents_cleaned <- rents_cleaned[-c(1, 2), ] # Remove the original rows 3 and 4 used for headers

# Correct column names
colnames(rents_cleaned) <- c(
  "Code",
  "Area",
  "2011 Q2", "2011 Q3", "2011 Q4",
  "2012 Q1", "2012 Q2", "2012 Q3", "2012 Q4",
  "2013 Q1", "2013 Q3",
  "2014 Q1", "2014 Q3",
  "2015 Q1", "2015 Q3",
  "2016 Q1", "2016 Q3",
  "2017 Q1", "2017 Q3",
  "2018 Q1", "2018 Q3",
  "2019 Q1"
)

rents_cleaned <- rents_cleaned %>%
  mutate(across(starts_with("201"), ~ as.numeric(.)))

# Restructure dataset by calculating the total for each year
rents_cleaned_annual <- rents_cleaned %>%
  rowwise() %>%
  mutate(
    `2011` = sum(c_across(starts_with("2011 Q")), na.rm = TRUE),
    `2012` = sum(c_across(starts_with("2012 Q")), na.rm = TRUE),
    `2013` = sum(c_across(starts_with("2013 Q")), na.rm = TRUE),
    `2014` = sum(c_across(starts_with("2014 Q")), na.rm = TRUE),
    `2015` = sum(c_across(starts_with("2015 Q")), na.rm = TRUE),
    `2016` = sum(c_across(starts_with("2016 Q")), na.rm = TRUE),
    `2017` = sum(c_across(starts_with("2017 Q")), na.rm = TRUE),
    `2018` = sum(c_across(starts_with("2018 Q")), na.rm = TRUE),
    `2019` = sum(c_across(starts_with("2019 Q")), na.rm = TRUE)
  ) %>%
  ungroup() %>%
  dplyr::select(Code, Area, `2011`:`2019`) # Keep only the summarized years

rents_final <- rents_cleaned_annual %>%
  slice(-c(34:39, 41:44))

# Rename columns
rents_final <- rents_final %>%
  rename(LAD11CD = Code, LAD11NM = Area)

# Perform a full join with the MSOA shapefile data
finals_rents <- full_join(msoa, rents_final, by = c("LAD11CD", "LAD11NM"))

# Keeping only specific columns
finals_rents <- finals_rents  %>%
  dplyr::select(MSOA11CD, MSOA11NM, geometry, Proportion, "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019")

rents_long <- finals_rents %>%
  pivot_longer(
    cols = starts_with("201"),  
    names_to = "Year",          
    values_to = "Rent"          
  ) %>%
  dplyr::select(MSOA11CD, MSOA11NM, geometry, Proportion, Year, Rent)

rents_long <- rents_long %>%
  mutate(
    Rent = ifelse(
      MSOA11CD == "London",
      Rent,                            
      Rent * Proportion                 
    )
  )

rents_long <- rents_long %>%
  dplyr::select(-Proportion)

rents_long <- rents_long %>%
  mutate(Year = as.integer(Year)) 

# Define all the years you want to interpolate
all_years <- 2001:2021

# Perform interpolation for the Rent column
interpolated_rents <- rents_long %>%
  group_by(MSOA11CD, MSOA11NM) %>%
  complete(Year = all_years) %>%  
  arrange(Year) %>%
  mutate(
    Rent = if (sum(!is.na(Rent)) > 1) {
      approx(
        x = Year[!is.na(Rent)], 
        y = Rent[!is.na(Rent)], 
        xout = Year, 
        rule = 2
      )$y
    } else {
      Rent  
    }
  ) %>%
  ungroup()
```

## Final housing data
```{r}
# Filter datasets for odd years
datasets <- list(
  interpolated_rents,
  final_budget,
  interpolated_tenure,
  combined_houseprice,
  accommodation_with_london
)

filtered_datasets <- lapply(datasets, function(data) {
  data %>%
    mutate(Year = as.numeric(as.character(Year))) %>% # Ensure Year is numeric
    filter(Year %% 2 != 0)                            # Keep only odd years
})

# Assign filtered datasets back to their original names
interpolated_rents <- filtered_datasets[[1]]
final_budget <- filtered_datasets[[2]]
final_tenancy <- filtered_datasets[[3]]
combined_houseprice <- filtered_datasets[[4]]
accommodation_with_london <- filtered_datasets[[5]]

datasets_with_geometry <- list(
  interpolated_rents,
  final_budget,
  final_tenancy,
  combined_houseprice,
  accommodation_with_london
)

standardized_datasets <- lapply(datasets_with_geometry, function(data) {
  if (!inherits(data, "sf") && "geometry" %in% names(data)) {
    st_as_sf(data, sf_column_name = "geometry") # Convert to sf
  } else {
    data # Keep as is if already sf
  }
})

# Assign standardized datasets back to their names
interpolated_rents <- standardized_datasets[[1]]
final_budget <- standardized_datasets[[2]]
final_tenancy <- standardized_datasets[[3]]
combined_houseprice <- standardized_datasets[[4]]
accommodation_with_london <- standardized_datasets[[5]]

# Assign CRS-standardized datasets back to their names
interpolated_rents <- standardized_datasets[[1]]
final_budget <- standardized_datasets[[2]]
final_tenancy <- standardized_datasets[[3]]
combined_houseprice <- standardized_datasets[[4]]
accommodation_with_london <- standardized_datasets[[5]]

# Full join all datasets
joined_data <- interpolated_rents %>%
  st_drop_geometry() %>%
  left_join(st_drop_geometry(final_budget), by = c("MSOA11CD", "Year")) %>%
  left_join(st_drop_geometry(final_tenancy), by = c("MSOA11CD", "Year")) %>%
  left_join(st_drop_geometry(combined_houseprice), by = c("MSOA11CD", "Year")) %>%
  left_join(st_drop_geometry(accommodation_with_london), by = c("MSOA11CD", "Year"))

cleaned_housing <- joined_data %>%
  # Rename and select only the unique columns
  dplyr::select(
    MSOA11CD,
    MSOA11NM = MSOA11NM.x,
    Year,
    Rent,
    budget_proportion,
    average_owned,
    average_rented,
    average_social,
    Houseprice,
    Occupation
  )

# Add missing London data
manual_values <- tibble(
  MSOA11CD = "London",
  MSOA11NM = "London",
  Year = c(2001, 2003, 2005, 2007, 2009, 2011, 2013, 2015, 2017, 2019, 2021),
  average_owned = c(0.5557597, 0.5557597, 0.5557597, 0.5534036, 0.5322169, 0.5061974, 
                    0.5193679, 0.5167916, 0.5241732, 0.5064820, 0.5459837),
  average_rented = c(0.18802592, 0.18802592, 0.18802592, 0.19389581, 0.22340161, 
                     0.24892051, 0.24442112, 0.24791433, 0.25317208, 0.26583293, 0.24447031),
  average_social = c(0.25621433, 0.25621433, 0.25621433, 0.25270054, 0.24438150, 
                     0.24485105, 0.23621099, 0.23529412, 0.22265474, 0.22765647, 0.20954598)
)

# Update the relevant rows in cleaned_housing
cleaned_housing <- cleaned_housing %>%
  mutate(
    average_owned = ifelse(MSOA11CD == "London" & Year %in% manual_values$Year, 
                           manual_values$average_owned[match(Year, manual_values$Year)], 
                           average_owned),
    average_rented = ifelse(MSOA11CD == "London" & Year %in% manual_values$Year, 
                            manual_values$average_rented[match(Year, manual_values$Year)], 
                            average_rented),
    average_social = ifelse(MSOA11CD == "London" & Year %in% manual_values$Year, 
                            manual_values$average_social[match(Year, manual_values$Year)], 
                            average_social)
  )
```

```{r}
# Specify columns for which you want to create lagged-1 versions
columns_to_lag <- c("Rent", "budget_proportion", "Houseprice", "Occupation", "average_social")

# Create lagged-1 columns
cleaned_housing <- cleaned_housing %>%
  arrange(MSOA11CD, Year) %>% # Ensure data is sorted
  mutate(across(all_of(columns_to_lag), ~ lag(.x, 1), .names = "lagged_{.col}"))
```

## Download Housing CSV
```{r}
# Write the cleaned data frame to a CSV file
write.csv(cleaned_housing, "housing_data.csv", row.names = FALSE)
```

## Neighborhood data

```{r}
# Load the datasets

setwd("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/neighborhood data")

greenery_df <- read_csv("greenery boroughs.csv")
crime_df <- read_csv("historic crime 2010-2021.csv")
ptal2015 <- read_csv("lsoa 2015 ptal.csv")
ptal2021 <- read_csv("grid 2021 ptal.csv")

# Spatial data
msoa <- st_read("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/MSOA_2011_London_gen_MHW.shp")
lsoa <- st_read("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/LSOA_2011_London_gen_MHW.shp")
borough <- st_read("/Users/anamariasteinercorrea/Desktop/v3 longitudinal diaspora study/Data/London_Borough_Excluding_MHW.shp")
```

### Greenery

```{r}
greenery_df_cleaned <- greenery_df %>%
  rename(area = lb_code) %>%  
  dplyr::select(area, percent_green) %>%  
  na.omit() %>%  
  mutate(percent_green = as.numeric(percent_green))

# Create a data frame with years 
years <- seq(2001, 2021, by = 2)

# Use merge() to perform a Cartesian join
expanded_grenery <- merge(greenery_df_cleaned, years)

# Change column name
names(expanded_grenery)[names(expanded_grenery) == "y"] <- "Year"

# Join to msoa chp
borough_greenery <- expanded_grenery %>%
  left_join(msoa, by = c("area" = "LAD11NM"))

# Calculate greenery of each msoa
borough_pop <- borough_greenery %>%
  group_by(LAD11CD) %>%
  summarize(total_borough_population = sum(USUALRES))

# Merge total borough population back to the original dataframe
borough_greenery <- merge(borough_greenery, borough_pop, by = "LAD11CD")

# Calculate the proportion of each MSOA in the borough
borough_greenery$population_proportion <- borough_greenery$USUALRES / borough_greenery$total_borough_population

# Assign greenery percentage to each MSOA
borough_greenery$MSOA_percent_green <- borough_greenery$population_proportion * borough_greenery$percent_green

# Select relevant columns for the result
msoa_green <- dplyr::select(borough_greenery, MSOA11CD, MSOA11NM, Year, MSOA_percent_green)
```
### Crime

```{r}
# Rename 'Borough' column to 'Code'
crime_df <- crime_df %>% rename(code = Borough)

# Aggregate crime values by 'Code'
aggregated_crime <- crime_df %>%
  group_by(code) %>%
  summarise(across(starts_with("20"), sum, na.rm = TRUE))

# Sum values by year
aggregated_crime <- aggregated_crime %>%
  gather(key = "YearMonth", value = "Count", -code) %>%
  mutate(Year = substr(YearMonth, 1, 4)) %>%
  group_by(code, Year) %>%
  summarise(Yearly_Total = sum(Count)) %>%
  spread(key = Year, value = Yearly_Total)

# Filter to keep only years 2011-2018
years_to_keep <- as.character(seq(2011, 2018, by = 1))
odd_years <- years_to_keep[as.numeric(years_to_keep) %% 2 != 0]
final_crime <- dplyr::select(aggregated_crime, code, all_of(odd_years))

# Transform to long format
long_crime <- final_crime %>%
  pivot_longer(cols = `2011`:`2017`, names_to = "Year", values_to = "crime_counts")

# Define the range of odd years
full_years <- seq(2001, 2021, by = 2)

# Interpolation function
interpolate_crime_counts <- function(long_crime) {
  interpolated <- approx(
    x = long_crime$Year, 
    y = long_crime$crime_counts, 
    xout = full_years, 
    rule = 2  # use last observation for extrapolation
  )
  return(data.frame(Year = interpolated$x, crime_counts = interpolated$y))
}

# Apply interpolation for each 'code'
interpolated_crime <- long_crime %>%
  group_by(code) %>%
  do(interpolate_crime_counts(.)) %>%
  ungroup()

# Join to msoa chp
borough_crime <- interpolated_crime %>%
  left_join(msoa, by = c("code" = "LAD11CD"))

# Calculate greenery of each msoa
borough_pop <- borough_crime %>%
  group_by(code) %>%
  summarize(total_borough_population = sum(USUALRES))

# Merge total borough population back to the original dataframe
borough_crime <- merge(borough_crime, borough_pop, by = "code")

# Calculate the proportion of each MSOA in the borough
borough_crime$population_proportion <- borough_crime$USUALRES / borough_crime$total_borough_population

# Assign greenery percentage to each MSOA
borough_crime$MSOA_crime <- borough_crime$population_proportion * borough_crime$crime_counts

# Select relevant columns for the result
msoa_crime <- dplyr::select(borough_crime, MSOA11CD, MSOA11NM, Year, MSOA_crime)
```
### PTAL 

```{r}
# Convert ptal2021 to an sf object
ptal_2021_sf <- ptal2021 %>%
  st_as_sf(coords = c("X", "Y"), crs = 27700)

lsoa <- st_transform(lsoa, crs = 27700)

# Perform spatial join with LSOA layer
ptal_lsoa_joined <- st_join(ptal_2021_sf, lsoa, join = st_within)

# Join ptals
final_ptal <- ptal_lsoa_joined %>%
  left_join(ptal2015, by = c("LSOA11CD" = "LSOA2011"))

clean_ptal <- dplyr::select(final_ptal, MSOA11CD, MSOA11NM, AI2021, AvPTAI2015)

clean_ptal <- clean_ptal %>%
  rename(`2021` = AI2021, `2015` = AvPTAI2015)

ptal_no_geometry <- st_drop_geometry(clean_ptal)

msoa_ptal <- ptal_no_geometry %>%
  group_by(MSOA11CD, MSOA11NM) %>%
  summarize(
    avg_2021 = mean(`2021`, na.rm = TRUE),
    avg_2015 = mean(`2015`, na.rm = TRUE),
  )

# Interpolate
# Transform to long format
ptal_long <- msoa_ptal %>%
  pivot_longer(cols = starts_with("avg"), names_to = "Year", values_to = "PTAL") %>%
  mutate(Year = as.integer(gsub("avg_", "", Year)),
         PTAL = round(PTAL))

library(dplyr)

# Define the range of full years for interpolation
full_years <- seq(2001, 2021, by = 2)

# Interpolation function
interpolate_ptal <- function(ptal_long) {
  interpolated <- approx(
    x = ptal_long$Year, 
    y = ptal_long$PTAL, 
    xout = full_years, 
    rule = 2  # using last observation for extrapolation
  )
  return(data.frame(Year = interpolated$x, PTAL = interpolated$y))
}

# Filter out groups with fewer than 2 non-NA PTAL values
filtered_ptal_long <- ptal_long %>%
  group_by(MSOA11NM) %>%
  filter(sum(!is.na(PTAL)) >= 2) %>%
  ungroup()

# Apply interpolation for each 'MSOA11NM'
interpolated_ptal <- filtered_ptal_long %>%
  group_by(MSOA11NM) %>%
  do(interpolate_ptal(.)) %>%
  ungroup()
```

## Neighborhood data

```{r}
# Join neighborhood dataset
neighborhood <- msoa_crime %>%
  full_join(msoa_green, by = c("MSOA11NM", "Year")) %>%
  full_join(interpolated_ptal, by = c("MSOA11NM", "Year"))

neighborhood_clean <- neighborhood %>%
  drop_na()

names(neighborhood_clean)

library(dplyr)

# Correcting the code to remove a column and rename another
neighborhood_clean <- neighborhood_clean %>%
  dplyr::select(-`MSOA11CD.y`) %>%
  dplyr::rename(MSOA11CD = `MSOA11CD.x`)

final_neighborhood_unique <- neighborhood_clean%>%
  group_by(MSOA11CD, Year) %>%
  slice(1) %>%  # or use distinct() if all other columns are the same
  ungroup()

final_neighborhood_clean <- final_neighborhood_unique %>%
  drop_na()
```

## Download neighborhood CSV
```{r}
# Write the cleaned data frame to a CSV file
write.csv(final_neighborhood_clean, "neighborhood_data.csv", row.names = FALSE)
```


